### 2012-08-28 paper.tex
Added all the files

%\documentclass[draftcls, onecolumn, 11pt]{../../bib/IEEEtran}
\documentclass[journal,10pt]{IEEEtran}

\usepackage{mathbf-abbrevs}

\input{defs}

\begin{document}

\title{On the Cram\'{e}r-Rao bound for polynomial phase signals}

\author{Robby McKilliam and Andr\'{e} Pollok%
\thanks{Supported under the Australian Government's Australian Space Research Program.
Robby McKilliam and Andr\'{e} Pollok are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095}
}
% The paper headers
\markboth{Robby McKilliam and Andre Pollock, On the Cram\'{e}r-Rao bound for polynomial phase signals}%
{DRAFT \today}

% make the title area
\maketitle

\begin{abstract}
Polynomial-phase signals have numerous applications including radar, sonar, geophysics, and radio communication.  Of particular importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.  Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.  The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.  In this paper we derive closed form formula for the exact Cram\'{e}r-Rao bounds.  This is achieved by making use of a family of orthogonal polynomials.
\end{abstract}

\begin{IEEEkeywords}
Polynomial-phase signals, Cram\'{e}r-Rao bound.
\end{IEEEkeywords}

\section{Introduction}

A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991, Rihaczek1996},
\[
s_n(b) = \alpha \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
\]
where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients, $j = \sqrt{-1}$ and $\Delta$ is the interval between consecutive samples.  We will assume, without loss of generality, that $\Delta = 1$.

Polynomial-phase signals have numerous applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000}.  The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995}.  Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)$ from $N$ consecutive samples $y_1, \dots. y_N$ where 
\[
y_n = s_n(\beta) + w_n
\]
and $\{w_n, n \in \ints\}$ is a sequence of noise variables.  Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994}.

Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.  Their approach requires the inversion of a poorly conditioned Fisher information matrix.  To avoid this, they provide an approximation that is valid when the number of observations $N$ is much larger than the order of the polynomial phase signal $m$.  Here we remove the need for this approximation.  By converting the polynomial basis to one that is orthogonal we derive closed form formula for the inverse Fisher information matrix.  This leads to closed form formula for the Cram\'{e}r-Rao bounds.  %At the same time we tie together some existing literature on the Cram\'{e}r-Roa bound for polynomial phase signals~\cite{Peleg1991_CRB_PPS_1991,Ristic_crbpelegcomment_1998}.

The paper is organised as follows.  In Section~\ref{sec_crb} we derive the Fisher information matrix in a similar way to Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}.  The matrix is poorly conditioned and inverting it to obtain the Cram\'{e}r-Rao bound is difficult.  In Section~\ref{sec:changing-basis} we describe some general results regarding changing the basis in which a polynomial phase signal may be represented.  We describe the Fisher information matrix corresponding with a chosen basis and we relate the matrices between two bases.  In Section~\ref{sec:discr-orth-polyn} we introduce the family of discrete orthogonal polynomials.  By transforming into this basis, the corresponding Fisher information matrix becomes diagonal, making matrix inversion simple.  This leads directly to closed form expressions for the Cram\'{e}r-Rao bounds.  To keep our derivation simple we initially assume that the amplitude $\alpha$ is a fixed constant, and not a parameter to be estimated.  However, including the amplitude is straightforward, and we do this in Section~\ref{sec:including-amplitude}.  Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations, these are in agreement with our exact bounds.

\section{The Cram\'{e}r-Rao bound} \label{sec_crb}

We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element having independent real and imaginary parts with zero mean and variance $\sigma^2$.  The probability density function of $w_n$ is 
\[
p(x) = \frac{1}{4 \pi \sigma^2} \exp\left( -\frac{\abs{x}^2}{2\sigma^2} \right)
\]
where $\abs{x}^2$ is the squared magnitude of the complex number $x$.  We wish to estimate $\beta$ from $y_1, \dots, y_N$.  The likelihood function is
\[
L(b) =  \frac{1}{(4 \pi \sigma^2)^N} \exp\left( - \frac{1}{2\sigma^{2}} \sum_{n=1}^N \abs{y_n -  s_n(b) }^2 \right).
\]
%The maximum likelihood estimators of $\alpha, \beta_0, \dots, \beta_m$ are the maximisers of $L$.
% , that is
% \[
% (\hat{\alpha}, \hat{\beta}_0, \dots, \hat{\beta}_m) = \arg\max_{a \in \reals, (b_0, \dots, b_m) \in B} L(a, b_0, \dots, b_m) 
% \]  
% where $B \subset \reals^{m+1}$ describes a region of polynomial phase coefficients for which the estimation problem is identifiable~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_LSU_polyest_part1_2012}.  We assume that $(\beta_0, \dots, \beta_m) \in B$.
For $i,k \in \{0, \dots, m\}$ define functions 
\[
f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}}.
\]  
The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
\[
F_{ik} = F_{ki} = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} n^{i}n^{k}.
\]

Let $\hat{\beta}$ be an unbiased estimator of $\beta$.  The Cram\'{e}r-Rao bound asserts:  If $C$ is the covariance matrix of $\hat{\beta}$, then $C - F^{-1}$ is positive semidefinite where $F^{-1}$ is the inverse of $F$.  The diagonal elements of a positive semidefinite matrix are nonnegative so $C_{ii} \geq F^{-1}_{ii}$ and correspondingly, 
\[
C_{ii} = \var \hat{\beta}_{i} \geq F_{ii}^{-1}, \qquad i=0,\dots,m.
\]  
Be aware that $F_{ik}^{-1}$ is the $ik$th element of the inverse matrix $F^{-1}$.  It is not the reciprocal of $F_{ik}$.  Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} found that $F$ is poorly conditioned and for this reason $F^{-1}$ is difficult to compute.  They give an approximation that is accurate when $N$ is sufficiently larger than $m$.  We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.  This will lead to closed form formula for $F^{-1}$.

\section{Changing the basis}\label{sec:changing-basis}

\newcommand{\calP}{{\mathcal P}}

For a polynomial $g(x) = b_0 + b_1 x + \dots + b_m x^m$ of order at most $m$ denote by $\coef(g) = (b_0, b_1, \dots, b_m)$ the vector of length $m+1$ containing its coefficients.  Let $p_0, \dots, p_m$ be a family of linearly independent polynomials.  By linearly independent it is meant that any polynomial of order at most $m$ can be written as a linear combination $c_0p_0 + c_1 p_1 + \dots + c_m p_m$ where $c_1, \dots, c_m$ are real numbers.  Observe that $p_0, \dots, p_m$ are linearly independent if and only if the vectors $\coef(p_0), \dots, \coef(p_m)$ are linearly independent.  We call $p_0, \dots, p_m$ a \emph{basis} for the space of polynomials of order $m$.  We call the polynomials $n^0, n^1, \dots, n^m$ the \emph{standard basis}.

Let 
\[
P = \left( \begin{array}{c}
\coef(p_0) \\
\vdots \\
\coef(p_m)
\end{array} \right)
\] 
be the $m+1$ by $m+1$ matrix with $k$th row $\coef(p_k)$.  If
\[
b_0 + b_1 x + \dots + b_m x^m = c_0 p_0(x) + \dots + c_m p_m(x)
\]
then the relationship between the vectors $b =  (b_0, \dots, b_m)$ and $c =  (c_0, \dots, c_m)$ is $b = cP$.  Since $P$ is invertible we also have $c = bP^{-1}$.

Put,
\[
r_n(c) = \alpha \exp\left( j \sum_{k = 0}^{m}{c_k p_k(n)}\right)
\]
so that $r_n(c)$ is a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.  Correspondingly $s_n(b)$ is a polynomial phase expressed in terms of the standard basis $n^0, n^1, \dots, n^m$.  The two signals are related by
\[
r_n(c) = s_n(cP) = s_n(b).
\]
The observed samples $y_1,\dots,y_N$ can be written using $r_n$ as
\[
y_n = r_n(\gamma) + w_n = s_n(\gamma P) + w_n
\]
where $\beta = \gamma P$.  We can now consider the problem of estimating $\gamma$ instead of $\beta$.  If $\hat{\beta}$ is an estimator of $\beta$ then the corresponding estimator of $\gamma$ is $\hat{\gamma} = \hat{\beta}P^{-1}$.  If $C$ is the covariance of $\hat{\beta}$ then the covariance of $\hat{\gamma}$ is $(P^{-1})^\prime C P^{-1}$ where $^\prime$ is the transpose.

The likelihood function for $\gamma$ is 
\[
L_\gamma(c) = L(cP) = L(b)
\] 
and we can compute the corresponding Fisher information matrix.  It is found to be the $m+1$ by $m+1$ matrix $H$ with elements
\[
H_{ik} = H_{ki} = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} p_{i}(n) p_{k}(n).
\]
The Cram\'{e}r-Rao bound for $\gamma$ is found by taking the inverse of $H$.  The next lemma relates the two Fisher information matrices $H$ and $F$.

\begin{lemma}
Suppose we desire to estimate $\beta \in \reals^m$ and that $L(b)$ is the corresponding likelihood function and $F$ the corresponding $m \times m$ Fisher information matrix.  Let $P$ be an $m \times m$ invertible matrix and $\gamma \in \reals^m$ be such that $\beta = \gamma P$.  Then the likelihood function corresponding to $\gamma$ is $L_\gamma(c) = L(cP)$ and the corresponding Fisher information matrix is $H = P F P^\prime$. 
\end{lemma}
\begin{IEEEproof}
Put $b = cP$.  That $L_\gamma(c) = L(cP) = L(b)$ is plain.  The elements $F_{ik} = f_{ik}(\beta)$ where
\[
f_{ik}(b) = -\expect \frac{\partial \log L(b)}{\partial b_i \partial b_k}.
\]
The elements $H_{ik} = h_{ik}(\gamma)$ where
\[
h_{ik}(c) = -\expect \frac{\partial \log L_\gamma(c)}{\partial c_i \partial c_k} =  -\expect \frac{\partial \log L(cP)}{\partial c_i \partial c_k}.
\]
Since $\frac{\partial b_s}{\partial c_i} = P_{is}$ applying the chain rule yields
\begin{align*}
h_{ik}(c) &= -\expect \sum_{s=0}^{m-1} P_{is}\frac{\partial \log L(cP)}{\partial b_s \partial c_k} \\
&= -\expect \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}   \frac{\partial \log L(b)}{\partial b_s \partial b_t} \\
&= \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  f_{st}(b).
\end{align*}
Putting $c = \gamma$ so that $b = \gamma P = \beta$ we have
\[
H_{ik} = \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  F_{st},
\]
or equivalently $H = P F P^\prime$.
\end{IEEEproof}

The lemma uses the notation we have developed for polynomial phase signals, however it is clear that the lemma applies generally, to any estimation problem for which a likelihood function and Fisher information matrix exist.  The utility of this result is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but, with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).  Using the lemma we have 
\[
F^{-1} = P^\prime H^{-1} P.
\]  
We will now apply this approach to compute the inverse Fisher information matrix for polynomial phase signals.  The most useful change of basis is described by the following family of polynomials.

\section{The discrete orthogonal polynomials}\label{sec:discr-orth-polyn}

\begin{definition}  \label{def:discreteLegendepolys}
The discrete orthogonal polynomial of order $k$, denoted by $d_k$, is
\[
d_k(x) = \frac{k!}{\binom{2k}{k}}\sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_s(x)},
\]
where $q_s$ is the polynomial
\[
q_s(x) = \binom{x-1}{s} = \frac{(x-1)(x-2)\dots(x-s)}{s!}
\]
and we define $q_0(x) = 1$.
\end{definition}
The discrete orthogonal polynomials (as we have defined them) are monic, i.e. the coefficient of the highest order term is equal to one, and the $k$th discrete orthogonal polynomial $d_k$ has order $k$.  The $d_k$ are orthogonal in the sense that~\cite{Szego1975_ortho_polynomials,Eisinberg2001_Vandermonde_rect,Eisinberg2007_discerete_otho_poly_equidist},
\[
\sum_{n=1}^{N}{ d_i(n) d_k(n) } = \begin{cases}
0 &  k\neq i \\
(k!)^2 \binom{2k}{k}^{-1} \binom{N+k}{2k+1}  & k = i,
\end{cases}
\]
and this will be a useful property for our purpose.  Let 
\[
P = \left( \begin{array}{c}
\coef(d_0) \\
\vdots \\
\coef(d_m)
\end{array} \right)
\]
and put $\beta = \gamma P$.  The Fisher information matrix corresponding to $\gamma$ is the $m+1$ by $m+1$ matrix $H$ with elements
\begin{align*}
H_{ik} &= \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} d_{i}(n) d_{k}(n) \\
&= \begin{cases}
0 &  k\neq i \\
 \frac{\alpha^2}{\sigma^2}\ (k!)^2 \binom{2k}{k}^{-1} \binom{N+k}{2k+1}  & k = i.
\end{cases}
\end{align*}
Thus $H$ is diagonal and the inverse has elements
\[
H_{ik}^{-1} = \begin{cases}
0 &  k\neq i \\
 \frac{\sigma^2}{\alpha^2}\ (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1}  & k = i.
\end{cases}
\] 
The inverse Fisher information matrix in the standard basis is $F^{-1} = P^\prime H^{-1} P$.  It remains to find closed form expressions for the elements of $P$, i.e. for the coefficients of $d_k$.

Observe that the polynomials $q_0,q_1,\dots$ satisfy the recursion
\[
q_s(x) = \frac{x-s}{s} q_{s-1}(x), \qquad q_{0}(x) = 1.
\]
If $q_s(x) = q_{s0} + q_{s1}x + \dots, q_{ss}x^s$ so that $q_{s0},q_{s1},\dots,q_{ss}$ are the coefficients of $q_{s}$, then
\[
q_{si} = \frac{1}{s}q_{s-1,i-1} - q_{s-1,i}
\]
where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if either of $s$ or $i$ is negative.

Now the $i$th coefficient of $d_k$ is
\[
P_{ki} = \frac{k!}{\binom{2k}{k}}\sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_{si}}
\]  
which is closed form in the sense that the number of operations required to compute $P_{ki}$ does not depend on $N$.  The inverse matrix $F^{-1} = P^\prime H^{-1} P$ can now be computed in closed form.  Evaluating the diagonal terms gives bounds,
\begin{align*}
\var \hat{\beta}_i \geq F_{ii}^{-1} &=  \sum_{s=0}^{m}\sum_{t=0}^{m} P_{si} P_{ti}  H_{st}^{-1} \\
&=  \sum_{k=0}^{m} P_{ki}^2  H_{kk}^{-1} \\
&= \frac{\sigma^2}{\alpha^2} \sum_{k=0}^{m} (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1} P_{ki}^2.
\end{align*}

%\section{Bounds derived from time shifting}

\section{Including the amplitude}\label{sec:including-amplitude}

Including the amplitude as an estimation parameter and deriving its Cram\'{e}r-Rao bound is straightforward.  Our model for a polynomial phase signal is now
\[
s_n(b) = a \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
\]
where $b=(b_0, \dots, b_m, a)$ is a vector of length $m+2$, the first $m+1$ elements being the polynomial coefficient (as before) and the last element being the amplitude.  We now consider estimation of $\beta = (\beta_0, \dots, \beta_m, \alpha)$ from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta) + w_n$.  The Fisher information matrix for $\beta$ can be shown to take the form
\[
G = \left( \begin{array}{cc}
F & \vdots \\
\cdots & \frac{N\alpha^2}{\sigma^2}
\end{array}\right)
\]
where $\cdots$'s indicate zeros.  The inverse is
\[
G^{-1} = \left( \begin{array}{cc}
F^{-1} & \vdots \\
\cdots & \frac{\sigma^2}{N\alpha^2}
\end{array} \right).
\]
So, the Cram\'{e}r-Rao bounds for the polynomial phase parameters $\beta_0, \dots, \beta_m$ are unchanged by including the amplitude.  The bound for an unbiased amplitude estimator $\hat{\alpha}$ is
\[
\var \hat{\alpha} \geq G_{m+1,m+1}^{-1} = \frac{\sigma^2}{N\alpha^2}.
\]
This is in agreement with Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}.

\section{Simulations}\label{sec:simuations}

Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} display the results of Monte-Carlo simulation with the maximum likelihood estimator and also the exact Cram\'{e}r-Rao bound and the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.  The simulations use $m=2$ and $N=5,20$ and with $\text{SNR} = \tfrac{\alpha^2}{2\sigma^2}$ in the range \unit[-5]{dB} to \unit[25]{dB}.   The average square error of the maximum likelihood estimator after $10^4$ replacations is given by the dots.  This is computed using the approach from~\cite[Sec.~10.1]{McKilliam2010thesis}.  The solid line is the exact Cram\'{e}r-Rao bound and the dashed line is the approximation.  As is expected, the maximum likelihood estimator achieves the Cram\'{e}r-Rao bound provided that the SNR is large enough, so that the `threshold' effect is avoided.  The approximate bound from~\cite{Peleg1991_CRB_PPS_1991} becomes accurate as $N$ increases.  In the figures the maximum likelihood estimator appears to perform better than the bound when the SNR is small.  This is an artifact of the identifiability conditions that must be imposed on the polynomial phase estimation problem~\cite[Ch.~7]{McKilliam2010thesis}\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.  The effect is that as SNR decreases, the variance of $\hat{\beta}_k$ converges to that of the uniform distribution on $[-\tfrac{0.5}{k!},\tfrac{0.5}{k!})$.

\begin{figure}[p]
   	\centering 
  		\includegraphics{code/plot-1.mps} 
   		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
   		\label{plot:phase0} 
  \end{figure}

\begin{figure}[p]
   	\centering 
  		\includegraphics{code/plot-2.mps} 
   		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
   		\label{plot:phase1} 
  \end{figure}

\begin{figure}[p]
   	\centering 
  		\includegraphics{code/plot-3.mps} 
   		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
   		\label{plot:phase2} 
  \end{figure}

\section{Conclusion}

Using a family of discrete orthogonal polynomials we provide closed form formula for the Cram\'{e}r-Rao lower bound for unbiased estimators of polynomial phase signals.

%\bibliographystyle{IEEEbib}
\small 
\bibliography{bib}

\end{document}

### 2012-08-28 build.sh

@@ -1,0 +1,9 @@
+cd code 
+mpost -interaction=nonstopmode plot.mp 
+cd .. 
+
+pdflatex paper.tex
+bibtex paper.aux
+pdflatex paper.tex
+pdflatex paper.tex
+

### 2012-08-28 code/sim.scala

@@ -1,0 +1,70 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+import pubsim.poly.PolynomialPhaseSignal
+import pubsim.poly.MaximumLikelihood
+import pubsim.distributions.GaussianNoise
+import pubsim.poly.PolynomialPhaseEstimator
+
+val iters = 10000 //number of Monte-Carlo trials.
+val Ns = List(5,20) //values of N we will generate curves for
+val ms = List(2) //order of our polynomial phase signals
+val snrdbs = -5 to 25
+val fourpisqr = 4.0*scala.math.Pi*scala.math.Pi
+
+
+//returns an array of noise distributions with a logarithmic scale
+def noises = snrdbs.map( snrdb => new GaussianNoise(0,scala.math.pow(10.0,-snrdb/10.0)/2.0) )
+
+//Return estimators we will run (factory patern to enable parallelism)
+def estfactory(m : Int, N : Int) : () => PolynomialPhaseEstimator = {
+    () => new MaximumLikelihood(m,N)
+}
+
+val starttime = (new java.util.Date).getTime
+
+//for all the the values of N and m.
+for( N <-  Ns; m <- ms ) {
+
+  val estf = estfactory(m,N)
+    
+    val estname = "simN" + N.toString + "m" + m
+    print("Running " + estname)
+    val eststarttime = (new java.util.Date).getTime
+    
+    //for all the noise distributions (in parallel threads)
+    val mselist = noises.par.map { noise =>
+
+      val siggen =  new PolynomialPhaseSignal(N) //construct a signal generator
+      siggen.setNoiseGenerator(noise)
+      val est = estf() //construct an estimator
+				  
+      var mse = new Array[Double](m+1) //storage for the mses
+      for( itr <- 1 to iters ) {
+	siggen.generateRandomParameters(m)
+	val p0 = siggen.getParameters
+	siggen.generateReceivedSignal
+	val err = est.error(siggen.getReal, siggen.getImag, p0)
+	for( i <- mse.indices ) mse(i) += err(i) 
+      }
+      print(".")
+      mse //last thing is what gets returned 
+    }.toList
+    
+    //now write all the data to a file
+    val files = (0 to m).map( v => new java.io.FileWriter(estname + "p" + v) ) //list of files to write to
+    (mselist, snrdbs).zipped.foreach{ (mse, snrdb) =>
[…]

### 2012-09-03 paper.tex
minor

@@
 In Section~\ref{sec:changing-basis} we describe some general results regarding changing the basis in which a polynomial phase signal may be represented.
-We describe the Fisher information matrix corresponding with a chosen basis and we relate the matrices between two bases.
+We describe the Fisher information matrix corresponding with a chosen basis, and we relate the matrices between two bases.
 In Section~\ref{sec:discr-orth-polyn} we introduce the family of discrete orthogonal polynomials.
@@
 However, including the amplitude is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
-Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations, these are in agreement with our exact bounds.
+Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations that agree with our exact bounds.
 
@@
 \[
-p(x) = \frac{1}{4 \pi \sigma^2} \exp\left( -\frac{\abs{x}^2}{2\sigma^2} \right)
+p(x) = \frac{1}{2 \pi \sigma^2} \exp\left( -\frac{\abs{x}^2}{2\sigma^2} \right)
 \]
@@
 \[
-L(b) =  \frac{1}{(4 \pi \sigma^2)^N} \exp\left( - \frac{1}{2\sigma^{2}} \sum_{n=1}^N \abs{y_n -  s_n(b) }^2 \right).
+L(b) = \frac{1}{(2 \pi \sigma^2)^N} \exp\left( - \frac{1}{2\sigma^{2}} \sum_{n=1}^N \abs{y_n -  s_n(b) }^2 \right).
 \]
@@
 Let $\hat{\beta}$ be an unbiased estimator of $\beta$.
-The Cram\'{e}r-Rao bound asserts:  If $C$ is the covariance matrix of $\hat{\beta}$, then $C - F^{-1}$ is positive semidefinite where $F^{-1}$ is the inverse of $F$.
+The Cram\'{e}r-Rao bound asserts:  If $C$ is the covariance matrix of $\hat{\beta}$, then $C - F^{-1}$ is positive semidefinite, where $F^{-1}$ is the inverse of $F$.
 The diagonal elements of a positive semidefinite matrix are nonnegative so $C_{ii} \geq F^{-1}_{ii}$ and correspondingly,

### 2012-10-17 paper.tex
minor

--- 2012-08-28 paper.tex
@@
 \]
-where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients, $j = \sqrt{-1}$ and $\Delta$ is the interval between consecutive samples.
+where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients (each a real number),  and $\Delta$ is the interval between consecutive samples.
 We will assume, without loss of generality, that $\Delta = 1$.
@@
 For a polynomial $g(x) = b_0 + b_1 x + \dots + b_m x^m$ of order at most $m$ denote by $\coef(g) = (b_0, b_1, \dots, b_m)$ the vector of length $m+1$ containing its coefficients.
-Let $p_0, \dots, p_m$ be a family of linearly independent polynomials.
+Let $p_0,\dots, p_m$ be a family of linearly independent polynomials of orders at most $m$.
 By linearly independent it is meant that any polynomial of order at most $m$ can be written as a linear combination $c_0p_0 + c_1 p_1 + \dots + c_m p_m$ where $c_1, \dots, c_m$ are real numbers.
@@
 We call $p_0, \dots, p_m$ a \emph{basis} for the space of polynomials of order $m$.
-We call the polynomials $n^0, n^1, \dots, n^m$ the \emph{standard basis}.
+We call the polynomials $x^0, x^1, \dots, x^m$ the \emph{standard basis}.
 
@@
 so that $r_n(c)$ is a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.
-Correspondingly $s_n(b)$ is a polynomial phase expressed in terms of the standard basis $n^0, n^1, \dots, n^m$.
+Correspondingly $s_n(b)$ is a polynomial phase signal expressed in terms of the standard basis $x^0, x^1, \dots, x^m$.
 The two signals are related by
@@
 \]
-where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if either of $s$ or $i$ is negative.
+where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if any of $s$ or $i$ is negative.
 

### 2012-10-17 paper.tex
added to paper and defs

@@
 \begin{abstract}
-Polynomial-phase signals have numerous applications including radar, sonar, geophysics, and radio communication.
+Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication.
 Of particular importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
@@
 \]
-where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), $b=(b_0, \dots, b_m)$ is a vector containing the polynomial coefficients (each a real number),  and $\Delta$ is the interval between consecutive samples.
+where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)$ is a vector representing the polynomial coefficients (each a real number),  and $\Delta$ is the interval between consecutive samples.
 We will assume, without loss of generality, that $\Delta = 1$.
 
-Polynomial-phase signals have numerous applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000}.
+Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000}.
 The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995}.
@@
 
-Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
-Their approach requires the inversion of a poorly conditioned Fisher information matrix.
-To avoid this, they provide an approximation that is valid when the number of observations $N$ is much larger than the order of the polynomial phase signal $m$.
+Under the assumption that $\{w_n\}$ are additive white and Gaussian with variance $\sigma^2$, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
+Their approach requires inversion of a poorly conditioned Fisher information matrix.
+To avoid this, they provide an approximation that is valid when the number of observations $N$ is sufficiently larger than the order of the polynomial phase signal $m$.
 Here we remove the need for this approximation.
@@
 This leads directly to closed form expressions for the Cram\'{e}r-Rao bounds.
-To keep our derivation simple we initially assume that the amplitude $\alpha$ is a fixed constant, and not a parameter to be estimated.
-However, including the amplitude is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
+To keep our derivation simple we initially assume that the amplitude $\alpha$ and the noise variance $\sigma^2$ are fixed constants, and not parameters to be estimated.
+However, including the amplitude and noise variance is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
 Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations that agree with our exact bounds.
@@
 \[
-f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}}.
+f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}},
 \]  
-The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
+where $\expect$ denotes expectation.  The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
 \[
@@
 \]
+It is a convenient property of polynomial phase estimation that the Fisher information matrix does not depend on the value of the polynomial phase coefficients being estimated.
+That is, $F$ does not depend on $\beta$.
 
@@
 Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} found that $F$ is poorly conditioned and for this reason $F^{-1}$ is difficult to compute.
-They give an approximation that is accurate when $N$ is sufficiently larger than $m$.
 We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.
@@
 \]
-The Cram\'{e}r-Rao bound for $\gamma$ is found by taking the inverse of $H$.  The next lemma relates the two Fisher information matrices $H$ and $F$.
+The Cram\'{e}r-Rao bound for $\gamma$ is found by taking the inverse of $H$.
+The next well know (even trivial) property relates the two Fisher information matrices $H$ and $F$.
+We do not intend to claim any novelty by including a formal statement of this property.
+The same property is, for example described in \cite[]{Kay1993_stat_sig_est_theory}.
+We included a proof simply for completeness.
 
-\begin{lemma}
+\begin{property}
 Suppose we desire to estimate $\beta \in \reals^m$ and that $L(b)$ is the corresponding likelihood function and $F$ the corresponding $m \times m$ Fisher information matrix.
@@
 Then the likelihood function corresponding to $\gamma$ is $L_\gamma(c) = L(cP)$ and the corresponding Fisher information matrix is $H = P F P^\prime$.
-\end{lemma}
+\end{property}
 \begin{IEEEproof}

### 2012-10-18 paper.tex
modified paper.

@@
 \]
-where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)$ is a vector representing the polynomial coefficients (each a real number),  and $\Delta$ is the interval between consecutive samples.
+where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)$ is a vector representing the polynomial coefficients (each $b_k$ a real number),  and $\Delta$ is the interval between consecutive samples.
 We will assume, without loss of generality, that $\Delta = 1$.
@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994}.
+Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994,Francos_2dpolyest_1998}.
 
-Under the assumption that $\{w_n\}$ are additive white and Gaussian with variance $\sigma^2$, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
+Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
 Their approach requires inversion of a poorly conditioned Fisher information matrix.
@@
 This leads directly to closed form expressions for the Cram\'{e}r-Rao bounds.
-To keep our derivation simple we initially assume that the amplitude $\alpha$ and the noise variance $\sigma^2$ are fixed constants, and not parameters to be estimated.
+To keep our derivation simple we initially assume that the amplitude and the noise variance are fixed constants, and not parameters to be estimated.
 However, including the amplitude and noise variance is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
-Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations that agree with our exact bounds.
+%Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations that agree with our exact bounds.
 
@@
 We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element having independent real and imaginary parts with zero mean and variance $\sigma^2$.
+Thus $\expect\abs{w_n}^2 = 2\sigma^2$.
 The probability density function of $w_n$ is
@@
 The Cram\'{e}r-Rao bound for $\gamma$ is found by taking the inverse of $H$.
-The next well know (even trivial) property relates the two Fisher information matrices $H$ and $F$.
+The next well know property relates the two Fisher information matrices $H$ and $F$.
 We do not intend to claim any novelty by including a formal statement of this property.
-The same property is, for example described in \cite[]{Kay1993_stat_sig_est_theory}.
-We included a proof simply for completeness.
+The property follows, for example from the more general result in~\cite[p.
+45]{Kay1993_stat_sig_est_theory}.
+%We included a proof simply for completeness.
 
@@
 \end{property}
-\begin{IEEEproof}
-Put $b = cP$.  That $L_\gamma(c) = L(cP) = L(b)$ is plain.  The elements $F_{ik} = f_{ik}(\beta)$ where
-\[
-f_{ik}(b) = -\expect \frac{\partial \log L(b)}{\partial b_i \partial b_k}.
-\]
-The elements $H_{ik} = h_{ik}(\gamma)$ where
-\[
-h_{ik}(c) = -\expect \frac{\partial \log L_\gamma(c)}{\partial c_i \partial c_k} =  -\expect \frac{\partial \log L(cP)}{\partial c_i \partial c_k}.
-\]
-Since $\frac{\partial b_s}{\partial c_i} = P_{is}$ applying the chain rule yields
-\begin{align*}
-h_{ik}(c) &= -\expect \sum_{s=0}^{m-1} P_{is}\frac{\partial \log L(cP)}{\partial b_s \partial c_k} \\
-&= -\expect \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}   \frac{\partial \log L(b)}{\partial b_s \partial b_t} \\
-&= \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  f_{st}(b).
-\end{align*}
-Putting $c = \gamma$ so that $b = \gamma P = \beta$ we have
-\[
-H_{ik} = \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  F_{st},
-\]
-or equivalently $H = P F P^\prime$.
-\end{IEEEproof}
+% \begin{IEEEproof}
+% Put $b = cP$.  That $L_\gamma(c) = L(cP) = L(b)$ is plain.  The elements $F_{ik} = f_{ik}(\beta)$ where
+% \[
+% f_{ik}(b) = -\expect \frac{\partial \log L(b)}{\partial b_i \partial b_k}.
+% \]
+% The elements $H_{ik} = h_{ik}(\gamma)$ where
+% \[
+% h_{ik}(c) = -\expect \frac{\partial \log L_\gamma(c)}{\partial c_i \partial c_k} =  -\expect \frac{\partial \log L(cP)}{\partial c_i \partial c_k}.
+% \]
+% Since $\frac{\partial b_s}{\partial c_i} = P_{is}$ applying the chain rule yields
+% \begin{align*}
+% h_{ik}(c) &= -\expect \sum_{s=0}^{m-1} P_{is}\frac{\partial \log L(cP)}{\partial b_s \partial c_k} \\
+% &= -\expect \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}   \frac{\partial \log L(b)}{\partial b_s \partial b_t} \\
+% &= \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  f_{st}(b).
+% \end{align*}
+% Putting $c = \gamma$ so that $b = \gamma P = \beta$ we have
+% \[
+% H_{ik} = \sum_{s=0}^{m-1}\sum_{t=0}^{m-1} P_{is} P_{kt}  F_{st},
+% \]
+% or equivalently $H = P F P^\prime$.
+% \end{IEEEproof}
 
-The lemma uses the notation we have developed for polynomial phase signals, however it is clear that the lemma applies generally, to any estimation problem for which a likelihood function and Fisher information matrix exist.
-The utility of this result is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but, with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).
-Using the lemma we have
+%The statement uses the notation we have developed for polynomial phase signals, but it applies generally, to any estimation problem for which a likelihood function and Fisher information matrix exist.
+The utility of this property is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but, with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).
+Using the property we have
 \[
@@
 the coefficient of the highest order term is equal to one, and the $k$th discrete orthogonal polynomial $d_k$ has order $k$.
-The $d_k$ are orthogonal in the sense that~\cite{Szego1975_ortho_polynomials,Eisinberg2001_Vandermonde_rect,Eisinberg2007_discerete_otho_poly_equidist},
+This family of polynomials is closely related to what are called the \emph{Gram polynomials} or \emph{discrete Chebyschev polynomials}~\cite[p.
+323]{Bjork_num_methods_least_square_1996}.
+The specific family given by Definition~\ref{def:discreteLegendepolys} also appears in~\cite{Eisinberg2007_discerete_otho_poly_equidist}.
+In this paper, we refer to the $d_k$ simply as the discrete orthogonal polynomials, but the reader should be aware that many other families of polynomials also go by this name.
+
+The $d_k$ are orthogonal in the sense that~\cite{Eisinberg2007_discerete_otho_poly_equidist},
 \[
@@
 
-\section{Including the amplitude}\label{sec:including-amplitude}
+\section{Including the amplitude and noise variance}\label{sec:including-amplitude}
 
-Including the amplitude as an estimation parameter and deriving its Cram\'{e}r-Rao bound is straightforward.  Our model for a polynomial phase signal is now
+Including the amplitude $\alpha$ and noise variance $\sigma^2$ as estimation parameters and deriving the corresponding Cram\'{e}r-Rao bounds is straightforward.
+Our model for a polynomial phase signal is now
 \[
-s_n(b) = a \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
+s_n(b,a) = a \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
 \]
-where $b=(b_0, \dots, b_m, a)$ is a vector of length $m+2$, the first $m+1$ elements being the polynomial coefficient (as before) and the last element being the amplitude.
-We now consider estimation of $\beta = (\beta_0, \dots, \beta_m, \alpha)$ from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta) + w_n$.
-The Fisher information matrix for $\beta$ can be shown to take the form
+where the argument $b=(b_0, \dots, b_m)$ represents the polynomial phase coefficients as before and the argument $a$ represents the amplitude.
+We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$, $\alpha$ and $\sigma^2$  from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta,\alpha) + w_n$.
+The likelihood function is now
 \[
-G = \left( \begin{array}{cc}
-F & \vdots \\
-\cdots & \frac{N\alpha^2}{\sigma^2}
+L(b,a,v) = \frac{1}{(2 \pi v )^N} \exp\left( - \frac{1}{2 v } \sum_{n=1}^N \abs{y_n -  s_n(b,a) }^2 \right).
+\]
+where the argument $v$ represents the noise variance.
+
+Pack the parameters $(\beta_0, \dots, \beta_m,\alpha,\sigma^2)$ into a vector of length $m+3$ and the corresponding Fisher information matrix can be shown to take the form
+\[
+G = \left( \begin{array}{ccc}
+F & \vdots & \vdots \\
+\cdots & \frac{N}{\sigma^2} & 0\\
+\cdots & 0 & \frac{N}{\sigma^4} \\
 \end{array}\right)
@@
 \[
-G^{-1} = \left( \begin{array}{cc}
-F^{-1} & \vdots \\
-\cdots & \frac{\sigma^2}{N\alpha^2}
+G^{-1} = \left( \begin{array}{ccc}
+F^{-1} & \vdots & \vdots \\
+\cdots & \frac{\sigma^2}{N} & 0 \\
+\cdots & 0 & \frac{\sigma^4}{N}
 \end{array} \right).
 \]
-So, the Cram\'{e}r-Rao bounds for the polynomial phase parameters $\beta_0, \dots, \beta_m$ are unchanged by including the amplitude.  The bound for an unbiased amplitude estimator $\hat{\alpha}$ is
+So, the Cram\'{e}r-Rao bounds for the polynomial phase coefficients $\beta_0, \dots, \beta_m$ are unchanged by including the amplitude and noise variance as parameters to be estimated.
+The bound for an unbiased variance estimator $\hat{\sigma}^2$ is
 \[
-\var \hat{\alpha} \geq G_{m+1,m+1}^{-1} = \frac{\sigma^2}{N\alpha^2}.
+\var \hat{\sigma}^2 \geq G_{m+2,m+2}^{-1} = \frac{\sigma^4}{N}.
+\]
+The bound for an unbiased amplitude estimator $\hat{\alpha}$ is
+\[
+\var \hat{\alpha} \geq G_{m+1,m+1}^{-1} = \frac{\sigma^2}{N}.
 \]
 This is in agreement with Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}.
+Note that in~\cite{Peleg1991_CRB_PPS_1991} the variance of the real and imaginary parts of $w_n$ is $\sigma^2/2$ rather than $\sigma^2$.
+This is explains the factor of $2$ difference between some of the formula in~\cite{Peleg1991_CRB_PPS_1991} and our own.
 
-\section{Simulations}\label{sec:simuations}
+% \section{Simulations}\label{sec:simuations}
 
-Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} display the results of Monte-Carlo simulation with the maximum likelihood estimator and also the exact Cram\'{e}r-Rao bound and the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.
+% Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} display the results of Monte-Carlo simulation with the maximum likelihood estimator and also the exact Cram\'{e}r-Rao bound and the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.
 The simulations use $m=2$ and $N=5,20$ and with $\text{SNR} = \tfrac{\alpha^2}{2\sigma^2}$ in the range \unit[-5]{dB} to \unit[25]{dB}.
@@
 
-\begin{figure}[p]
-   	\centering 
-  		\includegraphics{code/plot-1.mps} 
-   		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
-   		\label{plot:phase0} 
-  \end{figure}
+% \begin{figure}[p]
+%    	\centering 
+%   		\includegraphics{code/plot-1.mps} 
+%    		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
+%    		\label{plot:phase0} 
+%   \end{figure}
 
-\begin{figure}[p]
-   	\centering 
-  		\includegraphics{code/plot-2.mps} 
-   		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
-   		\label{plot:phase1} 
-  \end{figure}
+% \begin{figure}[p]
+%    	\centering 
+%   		\includegraphics{code/plot-2.mps} 
+%    		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
+%    		\label{plot:phase1} 
+%   \end{figure}
 
-\begin{figure}[p]
-   	\centering 
-  		\includegraphics{code/plot-3.mps} 
-   		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
-   		\label{plot:phase2} 
-  \end{figure}
+% \begin{figure}[p]
+%    	\centering 
+%   		\includegraphics{code/plot-3.mps} 
+%    		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
+%    		\label{plot:phase2} 
+%   \end{figure}
 

### 2012-10-18 paper.tex
more reviewer mods

@@
 
-A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991, Rihaczek1996},
+A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991},
 \[
@@
 \]
-where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)$ is a vector representing the polynomial coefficients (each $b_k$ a real number),  and $\Delta$ is the interval between consecutive samples.
+where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)\in\reals^{m+1}$ is a vector representing the polynomial coefficients,  and $\Delta$ is the interval between consecutive samples.
 We will assume, without loss of generality, that $\Delta = 1$.
@@
 The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995}.
-Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)$ from $N$ consecutive samples $y_1, \dots.
+Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)\in\reals^{m+1}$ from $N$ consecutive samples $y_1, \dots.
 y_N$ where
-\[
+\begin{equation}\label{eq:yndatamodel}
 y_n = s_n(\beta) + w_n
-\]
+\end{equation}
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
@@
 
-Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} have derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
+Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
 Their approach requires inversion of a poorly conditioned Fisher information matrix.
@@
 In Section~\ref{sec:changing-basis} we describe some general results regarding changing the basis in which a polynomial phase signal may be represented.
-We describe the Fisher information matrix corresponding with a chosen basis, and we relate the matrices between two bases.
+We describe the Fisher information matrix corresponding with a chosen basis, and relate the matrices between two bases.
 In Section~\ref{sec:discr-orth-polyn} we introduce the family of discrete orthogonal polynomials.
@@
 
-We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element having independent real and imaginary parts with zero mean and variance $\sigma^2$.
-Thus $\expect\abs{w_n}^2 = 2\sigma^2$.
+We make the assumption that the noise sequence $\{ w_n \}$ is independent and identically distributed complex Gaussian, each element $w_n$ having independent real and imaginary parts with zero mean and variance $\sigma^2$.
+Thus $\expect\abs{w_n}^2 = 2\sigma^2$ where $\expect$ denotes expectation.
 The probability density function of $w_n$ is
@@
 \[
-f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}},
+f_{ik}(b) = f_{ki}(b) = -\expect \frac{ \partial^2 \log L(b) }{ \partial b_{i} \partial b_{k}}.
 \]  
-where $\expect$ denotes expectation.  The Fisher information matrix $F$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
+The Fisher information matrix $F$ corresponding with $\beta$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
 \[
@@
 It is not the reciprocal of $F_{ik}$.
-Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} found that $F$ is poorly conditioned and for this reason $F^{-1}$ is difficult to compute.
+Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} found that $F$ can be poorly conditioned and for this reason $F^{-1}$ can be difficult to compute.
 We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.
@@
 \]
-so that $r_n(c)$ is a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.
-Correspondingly $s_n(b)$ is a polynomial phase signal expressed in terms of the standard basis $x^0, x^1, \dots, x^m$.
+so that $r_n(c)$ represents a polynomial phase signal expressed in terms of the basis $p_0, \dots, p_m$.
+Correspondingly $s_n(b)$ represents a polynomial phase signal expressed in terms of the standard basis $x^0, x^1, \dots, x^m$.
 The two signals are related by
@@
 \]
-The observed samples $y_1,\dots,y_N$ can be written using $r_n$ as
+The observed samples $y_1,\dots,y_N$ from~\eqref{eq:yndatamodel} can be written using $r_n$ as
 \[
@@
 The Cram\'{e}r-Rao bound for $\gamma$ is found by taking the inverse of $H$.
-The next well know property relates the two Fisher information matrices $H$ and $F$.
+It follows, for example from~\cite[Sec.
+3.8]{Kay1993_stat_sig_est_theory}, that the Fisher information matrices $H$ and $F$ are related by,
+\[
+H = P F P^\prime.
+\]
+The utility of this relationship is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).
+We then have
+\[
+F^{-1} = P^\prime H^{-1} P.
+\]  
+We will now apply this approach to compute the inverse Fisher information matrix for polynomial phase signals.  The most useful change of basis is described by the following family of polynomials.
+
+% The next well know property relates the two Fisher information matrices $H$ and $F$.
 We do not intend to claim any novelty by including a formal statement of this property.
@@
 
-\begin{property}
-Suppose we desire to estimate $\beta \in \reals^m$ and that $L(b)$ is the corresponding likelihood function and $F$ the corresponding $m \times m$ Fisher information matrix.
+% \begin{property}
+% Suppose we desire to estimate $\beta \in \reals^m$ and that $L(b)$ is the corresponding likelihood function and $F$ the corresponding $m \times m$ Fisher information matrix.
 Let $P$ be an $m \times m$ invertible matrix and $\gamma \in \reals^m$ be such that $\beta = \gamma P$.
 Then the likelihood function corresponding to $\gamma$ is $L_\gamma(c) = L(cP)$ and the corresponding Fisher information matrix is $H = P F P^\prime$.
-\end{property}
+% \end{property}
 % \begin{IEEEproof}
@@
 %The statement uses the notation we have developed for polynomial phase signals, but it applies generally, to any estimation problem for which a likelihood function and Fisher information matrix exist.
-The utility of this property is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but, with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).
+% The utility of this property is as follows:  Computation of the inverse matrix $F^{-1}$ may be difficult, but, with a carefully chosen change of basis matrix $P$, computation of $H^{-1}$ might be easy (for example $H$ could be diagonal).
 Using the property we have
-\[
-F^{-1} = P^\prime H^{-1} P.
-\]  
-We will now apply this approach to compute the inverse Fisher information matrix for polynomial phase signals.  The most useful change of basis is described by the following family of polynomials.
+% \[
+% F^{-1} = P^\prime H^{-1} P.
+% \]  
+% We will now apply this approach to compute the inverse Fisher information matrix for polynomial phase signals.  The most useful change of basis is described by the following family of polynomials.
 
@@
 This family of polynomials is closely related to what are called the \emph{Gram polynomials} or \emph{discrete Chebyschev polynomials}~\cite[p.
-323]{Bjork_num_methods_least_square_1996}.
+323]{Bjork_num_methods_least_square_1996}\cite{Chebyshev_discrete_polys1864,Gram_discrete_polys1883}.
 The specific family given by Definition~\ref{def:discreteLegendepolys} also appears in~\cite{Eisinberg2007_discerete_otho_poly_equidist}.
-In this paper, we refer to the $d_k$ simply as the discrete orthogonal polynomials, but the reader should be aware that many other families of polynomials also go by this name.
+In this paper, we refer to $d_0,d_1,\dots$ simply as the discrete orthogonal polynomials, but the reader should be aware that many other families of polynomials also go by this name.
 
@@
 \] 
-The inverse Fisher information matrix in the standard basis is $F^{-1} = P^\prime H^{-1} P$.  It remains to find closed form expressions for the elements of $P$, i.e. for the coefficients of $d_k$.
+The inverse Fisher information matrix in the standard basis is $F^{-1} = P^\prime H^{-1} P$.
+It remains to find closed form expressions for the elements of $P$, i.e.
+for the coefficients of $d_0,\dots,d_m$.
 
@@
 \]
-where the argument $b=(b_0, \dots, b_m)$ represents the polynomial phase coefficients as before and the argument $a$ represents the amplitude.
+where the argument $b=(b_0, \dots, b_m)$ represents the polynomial phase coefficients and the argument $a$ represents the amplitude.
 We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$, $\alpha$ and $\sigma^2$  from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta,\alpha) + w_n$.
@@
 Note that in~\cite{Peleg1991_CRB_PPS_1991} the variance of the real and imaginary parts of $w_n$ is $\sigma^2/2$ rather than $\sigma^2$.
-This is explains the factor of $2$ difference between some of the formula in~\cite{Peleg1991_CRB_PPS_1991} and our own.
+This explains the factor of $2$ difference between some of the formula in~\cite{Peleg1991_CRB_PPS_1991} and our own.
 

### 2012-10-19 paper.tex
edits to paper.

@@
 Of particular importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
-Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.
+%Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.
 The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.
-In this paper we derive closed form formula for the exact Cram\'{e}r-Rao bounds.
+In this paper we derive closed form formulae for the Cram\'{e}r-Rao bounds under the assumption that the noise is additive white and Gaussian.
 This is achieved by making use of a family of orthogonal polynomials.
@@
 A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991},
-\[
+\begin{equation}\label{eq:polyestfunction}
 s_n(b) = \alpha \exp\left( j \sum_{k = 0}^{m}{b_k (\Delta n)^k}\right)
-\]
+\end{equation}
 where $n \in \ints$, $\alpha$ is the signal amplitude (a positive real number), the argument $b=(b_0, \dots, b_m)\in\reals^{m+1}$ is a vector representing the polynomial coefficients,  and $\Delta$ is the interval between consecutive samples.
@@
 The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995}.
-Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)\in\reals^{m+1}$ from $N$ consecutive samples $y_1, \dots.
-y_N$ where
+Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)\in\reals^{m+1}$ from $N$ consecutive samples $y_1, \dots, y_N$ where
 \begin{equation}\label{eq:yndatamodel}
@@
 
+We remark on our notation.
+We use lower case Latin letters for working variables and we use Greek letters for fixed numbers.
+Where possible we use the Latin letter and its Greek counterpart when there is a relationship between a working variable and a fixed number.
+For example we use $b=(b_0,\dots,b_m)$ as a working variable to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
+It has been common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also the first element in the (working) vector $b$.
+
 Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
@@
 Here we remove the need for this approximation.
-By converting the polynomial basis to one that is orthogonal we derive closed form formula for the inverse Fisher information matrix.
-This leads to closed form formula for the Cram\'{e}r-Rao bounds.
+By converting the polynomial basis to one that is orthogonal we derive a closed form formula for the inverse Fisher information matrix.
+This leads to closed form formulae for the Cram\'{e}r-Rao bounds.
 %At the same time we tie together some existing literature on the Cram\'{e}r-Roa bound for polynomial phase signals~\cite{Peleg1991_CRB_PPS_1991,Ristic_crbpelegcomment_1998}.
@@
 We describe the Fisher information matrix corresponding with a chosen basis, and relate the matrices between two bases.
-In Section~\ref{sec:discr-orth-polyn} we introduce the family of discrete orthogonal polynomials.
+In Section~\ref{sec:discr-orth-polyn} we introduce a family of discrete orthogonal polynomials.
 By transforming into this basis, the corresponding Fisher information matrix becomes diagonal, making matrix inversion simple.
@@
 Let $\hat{\beta}$ be an unbiased estimator of $\beta$.
-The Cram\'{e}r-Rao bound asserts:  If $C$ is the covariance matrix of $\hat{\beta}$, then $C - F^{-1}$ is positive semidefinite, where $F^{-1}$ is the inverse of $F$.
+The Cram\'{e}r-Rao bound asserts:  If $C$ is the covariance matrix of $\hat{\beta}$, then $C - F^{-1}$ is positive semidefinite, where $F^{-1}$ is the inverse of $F$~\cite{Kay1993_stat_sig_est_theory}.
 The diagonal elements of a positive semidefinite matrix are nonnegative so $C_{ii} \geq F^{-1}_{ii}$ and correspondingly,
@@
 We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.
-This will lead to closed form formula for $F^{-1}$.
+This will lead to a closed form formula for $F^{-1}$.
 
@@
 &=  \sum_{k=0}^{m} P_{ki}^2  H_{kk}^{-1} \\
-&= \frac{\sigma^2}{\alpha^2} \sum_{k=0}^{m} (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1} P_{ki}^2.
+&= \frac{\sigma^2}{\alpha^2} \sum_{k=0}^{m} (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1} P_{ki}^2,
 \end{align*}
+where the second line follows because $H^{-1}$ is diagonal.
 
@@
 where the argument $b=(b_0, \dots, b_m)$ represents the polynomial phase coefficients and the argument $a$ represents the amplitude.
-We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$, $\alpha$ and $\sigma^2$  from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta,\alpha) + w_n$.
+We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$ and $\alpha$ and $\sigma^2$  from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta,\alpha) + w_n$.
 The likelihood function is now
@@
 Note that in~\cite{Peleg1991_CRB_PPS_1991} the variance of the real and imaginary parts of $w_n$ is $\sigma^2/2$ rather than $\sigma^2$.
-This explains the factor of $2$ difference between some of the formula in~\cite{Peleg1991_CRB_PPS_1991} and our own.
+This explains the factor of $2$ difference between some of the formulae in~\cite{Peleg1991_CRB_PPS_1991} and our own.
 
@@
 
-Using a family of discrete orthogonal polynomials we provide closed form formula for the Cram\'{e}r-Rao lower bound for unbiased estimators of polynomial phase signals.
+Using a family of discrete orthogonal polynomials we provide closed form formulae for the Cram\'{e}r-Rao lower bounds for unbiased estimators of polynomial phase signals.
 

### 2012-10-19 paper.tex
at resubmitted version

@@
 Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication.
-Of particular importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
+Of practical importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
 %Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.
 The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.
-In this paper we derive closed form formulae for the Cram\'{e}r-Rao bounds under the assumption that the noise is additive white and Gaussian.
+We derive closed form formulae for the Cram\'{e}r-Rao  lower bounds of these parameters under the assumption that the noise is additive white and Gaussian.
 This is achieved by making use of a family of orthogonal polynomials.
@@
 For example we use $b=(b_0,\dots,b_m)$ as a working variable to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
-It has been common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also the first element in the (working) vector $b$.
+It has been common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also an element in the vector $b$.
 
 Under the assumption that $\{w_n\}$ are additive white and Gaussian, Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991} derived the Cram\'{e}r-Rao lower bound for unbiased estimators of the coefficients $\beta$.
-Their approach requires inversion of a poorly conditioned Fisher information matrix.
+Their approach requires inversion of an $m+1$ by $m+1$ Fisher information matrix that can be poorly conditioned.
 To avoid this, they provide an approximation that is valid when the number of observations $N$ is sufficiently larger than the order of the polynomial phase signal $m$.
 Here we remove the need for this approximation.
-By converting the polynomial basis to one that is orthogonal we derive a closed form formula for the inverse Fisher information matrix.
+By converting the polynomial basis to one that is orthogonal we derive a closed form and numerically stable formula for the inverse Fisher information matrix.
 This leads to closed form formulae for the Cram\'{e}r-Rao bounds.
 %At the same time we tie together some existing literature on the Cram\'{e}r-Roa bound for polynomial phase signals~\cite{Peleg1991_CRB_PPS_1991,Ristic_crbpelegcomment_1998}.
+
+Strictly speaking, closed form formulae for the Cram\'{e}r-Rao bounds have been derived by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}, in the sense that inversion of an $m+1$ by $m+1$ matrix requires a number of arithmetic operations that does not depend on $N$.
+Nevertheless, we believe there is value in our closed form and numerically stable expressions, over, albeit closed form, but potentially numerically troublesome expressions involving the matrix inverse.
 
@@
 \]
-So, the Cram\'{e}r-Rao bounds for the polynomial phase coefficients $\beta_0, \dots, \beta_m$ are unchanged by including the amplitude and noise variance as parameters to be estimated.
+So, the Cram\'{e}r-Rao bounds for the polynomial phase coefficients $\beta_0, \dots, \beta_m$ are unchanged by considering the amplitude and noise variance as parameters to be estimated.
 The bound for an unbiased variance estimator $\hat{\sigma}^2$ is

### 2012-12-05 code/sim.scala
minor edits

@@ -8,7 +8,7 @@ import pubsim.distributions.GaussianNoise
 import pubsim.poly.PolynomialPhaseEstimator
 
 val iters = 10000 //number of Monte-Carlo trials.
-val Ns = List(5,20) //values of N we will generate curves for
+val Ns = List(10,20) //values of N we will generate curves for
 val ms = List(2) //order of our polynomial phase signals
 val snrdbs = -5 to 25
 val fourpisqr = 4.0*scala.math.Pi*scala.math.Pi

### 2012-12-06 paper.tex
minor edits
moved some stuff

@@
-%\documentclass[draftcls, onecolumn, 11pt]{../../bib/IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
 \documentclass[journal,10pt]{IEEEtran}
@@
 The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.
-We derive closed form formulae for the Cram\'{e}r-Rao  lower bounds of these parameters under the assumption that the noise is additive white and Gaussian.
+We derive closed form and numerically stable formulae for the Cram\'{e}r-Rao  lower bounds of these parameters under the assumption that the noise is additive white and Gaussian.
 This is achieved by making use of a family of orthogonal polynomials.
@@
 
-Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000}.
-The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995}.
+Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001,Hlawatsch_lin_quad_time_freq_spmag_1992}.
+The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
 Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)\in\reals^{m+1}$ from $N$ consecutive samples $y_1, \dots, y_N$ where
@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994,Francos_2dpolyest_1998}.
+Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994,Francos_2dpolyest_1998,McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
 
@@
 For example we use $b=(b_0,\dots,b_m)$ as a working variable to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
-It has been common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also an element in the vector $b$.
+It is common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also an element in the vector $b$.
 
@@
 By converting the polynomial basis to one that is orthogonal we derive a closed form and numerically stable formula for the inverse Fisher information matrix.
-This leads to closed form formulae for the Cram\'{e}r-Rao bounds.
+This leads to closed form and numerically stable formulae for the Cram\'{e}r-Rao bounds.
 %At the same time we tie together some existing literature on the Cram\'{e}r-Roa bound for polynomial phase signals~\cite{Peleg1991_CRB_PPS_1991,Ristic_crbpelegcomment_1998}.
 
-Strictly speaking, closed form formulae for the Cram\'{e}r-Rao bounds have been derived by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}, in the sense that inversion of an $m+1$ by $m+1$ matrix requires a number of arithmetic operations that does not depend on $N$.
+%Strictly speaking, closed form formulae for the Cram\'{e}r-Rao bounds have been derived by Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}, in the sense that inversion of an $m+1$ by $m+1$ matrix requires a number of arithmetic operations that does not depend on $N$.
 Nevertheless, we believe there is value in our closed form and numerically stable expressions, over, albeit closed form, but potentially numerically troublesome expressions involving the matrix inverse.
@@
 In Section~\ref{sec_crb} we derive the Fisher information matrix in a similar way to Peleg and Porat~\cite{Peleg1991_CRB_PPS_1991}.
-The matrix is poorly conditioned and inverting it to obtain the Cram\'{e}r-Rao bound is difficult.
+The matrix can be poorly conditioned and in this case inverting it to obtain the Cram\'{e}r-Rao bound is difficult.
 In Section~\ref{sec:changing-basis} we describe some general results regarding changing the basis in which a polynomial phase signal may be represented.
@@
 This leads directly to closed form expressions for the Cram\'{e}r-Rao bounds.
+In Section~\ref{sec:stab-cons} we manipulate these expression so as to make their numerical stability clear.
+This simultaneously indicates the prefered stable approach to implementing the formula in a digital computer.
 To keep our derivation simple we initially assume that the amplitude and the noise variance are fixed constants, and not parameters to be estimated.
 However, including the amplitude and noise variance is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
-%Section~\ref{sec:simuations} presents the results of Monte-Carlo simulations that agree with our exact bounds.
+Section~\ref{sec:simuations} SIMS BLERG.
 
@@
 We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.
-This will lead to a closed form formula for $F^{-1}$.
+This will lead to a closed form and numerically stable formula for $F^{-1}$.
 
@@
 \[
-d_k(x) = \frac{k!}{\binom{2k}{k}}\sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_s(x)},
+d_k(x) = \sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_s(x)},
 \]
@@
 \end{definition}
-The discrete orthogonal polynomials (as we have defined them) are monic, i.e.
+%The discrete orthogonal polynomials (as we have defined them) are monic, i.e.
 the coefficient of the highest order term is equal to one, and the $k$th discrete orthogonal polynomial $d_k$ has order $k$.
@@
 0 &  k\neq i \\
-(k!)^2 \binom{2k}{k}^{-1} \binom{N+k}{2k+1}  & k = i,
+\binom{2k}{k} \binom{N+k}{2k+1}  & k = i,
 \end{cases}
@@
 0 &  k\neq i \\
- \frac{\alpha^2}{\sigma^2}\ (k!)^2 \binom{2k}{k}^{-1} \binom{N+k}{2k+1}  & k = i.
+ \frac{\alpha^2}{\sigma^2} \binom{2k}{k} \binom{N+k}{2k+1}  & k = i.
 \end{cases}
@@
 0 &  k\neq i \\
- \frac{\sigma^2}{\alpha^2}\ (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1}  & k = i.
+ \frac{\sigma^2}{\alpha^2} \binom{2k}{k}^{-1} \binom{N+k}{2k+1}^{-1}  & k = i.
 \end{cases}
@@
 \[
-P_{ki} = \frac{k!}{\binom{2k}{k}}\sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_{si}}
+P_{ki} = \sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_{si}}
 \]  
@@
 &=  \sum_{k=0}^{m} P_{ki}^2  H_{kk}^{-1} \\
-&= \frac{\sigma^2}{\alpha^2} \sum_{k=0}^{m} (k!)^{-2} \binom{2k}{k} \binom{N+k}{2k+1}^{-1} P_{ki}^2,
+&= \frac{\sigma^2}{\alpha^2} \sum_{k=0}^{m} \binom{2k}{k}^{-1} \binom{N+k}{2k+1}^{-1} P_{ki}^2,
 \end{align*}
 where the second line follows because $H^{-1}$ is diagonal.
-
-%\section{Bounds derived from time shifting}
 

### 2012-12-06 code/crbcomputer.m
added new matlab and mathematica computers

@@ -1,0 +1,23 @@
+function computer
+
+% the q function, coefficients of the integer valued polynomials
+    function out = q(s,i)
+        if(s<0 || i <0) 
+            out = 0;
+        elseif(s==0 && i==0) 
+            out = 1;
+        elseif(s==0) 
+            out = 0;
+        else 
+            out = q(s-1,i-1)/s - q(s-1,i);
+        end
+    end
+    
+    %run some tests on q
+    if( q(0,0) ~= 1 ) 
+        disp('FAIL 1'); 
+    end
+    
+    
+    
+end

### 2012-12-06 code/crbcomputer.m
mved stuff

@@ -1,23 +1,0 @@
-function computer
-
-% the q function, coefficients of the integer valued polynomials
-    function out = q(s,i)
-        if(s<0 || i <0) 
-            out = 0;
-        elseif(s==0 && i==0) 
-            out = 1;
-        elseif(s==0) 
-            out = 0;
-        else 
-            out = q(s-1,i-1)/s - q(s-1,i);
-        end
-    end
-    
-    %run some tests on q
-    if( q(0,0) ~= 1 ) 
-        disp('FAIL 1'); 
-    end
-    
-    
-    
-end

### 2012-12-06 code/sim.scala

@@ -1,70 +1,0 @@
-/**
- * Run simulations for the sparse noisy period estimation problem.
- * Various estimators are available.
- */
-import pubsim.poly.PolynomialPhaseSignal
-import pubsim.poly.MaximumLikelihood
-import pubsim.distributions.GaussianNoise
-import pubsim.poly.PolynomialPhaseEstimator
-
-val iters = 10000 //number of Monte-Carlo trials.
-val Ns = List(10,20) //values of N we will generate curves for
-val ms = List(2) //order of our polynomial phase signals
-val snrdbs = -5 to 25
-val fourpisqr = 4.0*scala.math.Pi*scala.math.Pi
-
-
-//returns an array of noise distributions with a logarithmic scale
-def noises = snrdbs.map( snrdb => new GaussianNoise(0,scala.math.pow(10.0,-snrdb/10.0)/2.0) )
-
-//Return estimators we will run (factory patern to enable parallelism)
-def estfactory(m : Int, N : Int) : () => PolynomialPhaseEstimator = {
-    () => new MaximumLikelihood(m,N)
-}
-
-val starttime = (new java.util.Date).getTime
-
-//for all the the values of N and m.
-for( N <-  Ns; m <- ms ) {
-
-  val estf = estfactory(m,N)
-    
-    val estname = "simN" + N.toString + "m" + m
-    print("Running " + estname)
-    val eststarttime = (new java.util.Date).getTime
-    
-    //for all the noise distributions (in parallel threads)
-    val mselist = noises.par.map { noise =>
-
-      val siggen =  new PolynomialPhaseSignal(N) //construct a signal generator
-      siggen.setNoiseGenerator(noise)
-      val est = estf() //construct an estimator
-				  
-      var mse = new Array[Double](m+1) //storage for the mses
-      for( itr <- 1 to iters ) {
-	siggen.generateRandomParameters(m)
-	val p0 = siggen.getParameters
-	siggen.generateReceivedSignal
-	val err = est.error(siggen.getReal, siggen.getImag, p0)
-	for( i <- mse.indices ) mse(i) += err(i) 
-      }
-      print(".")
-      mse //last thing is what gets returned 
-    }.toList
-    
-    //now write all the data to a file
-    val files = (0 to m).map( v => new java.io.FileWriter(estname + "p" + v) ) //list of files to write to
-    (mselist, snrdbs).zipped.foreach{ (mse, snrdb) =>
[…]

### 2012-12-06 code/data/crbinversion.m
added data folder

@@ -1,0 +1,24 @@
+function vars = crbinversion(m,N,snrs)
+% computes the Cramer-Rao bound (CRB) for the polynomial phase
+% coefficients of a polynomial phase signal of order m given N
+% observations corrupted by additive white Gaussian noise.  The
+% vector snrs is a row vector with the desired values of signal to
+% noise ratio (SNR) to CRB at.  The output vars is an m+1 by length(snrs)
+% matrix such that vars(i,s) is the CRB for the coefficient of
+% order i when the SNR is equal to snrs(s).
+%
+% This attemps to invert the Fisher information matrix.  It's often
+% not numerically stable!
+
+%construct Vandermonde matrix
+V = zeros(m+1,N);
+for k=0:m
+    V(k+1,:)=(1:N).^k;
+end
+
+%construct Fisher information matrix
+F = V*V';
+
+vars = diag(inv(F)) * snrs;
+
+end

### 2012-12-06 code/data/sim.scala

@@ -1,0 +1,68 @@
+/**
+ * Run simulations for the sparse noisy period estimation problem.
+ * Various estimators are available.
+ */
+import pubsim.poly.PolynomialPhaseSignal
+import pubsim.poly.MaximumLikelihood
+import pubsim.distributions.GaussianNoise
+import pubsim.poly.PolynomialPhaseEstimator
+
+val iters = 10000 //number of Monte-Carlo trials.
+val Ns = List(10,20) //values of N we will generate curves for
+val ms = List(2) //order of our polynomial phase signals
+val snrdbs = -5 to 25
+val fourpisqr = 4.0*scala.math.Pi*scala.math.Pi
+
+//returns an array of noise distributions with a logarithmic scale
+def noises = snrdbs.map( snrdb => new GaussianNoise(0,scala.math.pow(10.0,-snrdb/10.0)/2.0) )
+
+//Return estimators we will run (factory patern to enable parallelism)
+def estfactory(m : Int, N : Int) : () => PolynomialPhaseEstimator = {
+    () => new MaximumLikelihood(m,N)
+}
+
+val starttime = (new java.util.Date).getTime
+
+//for all the the values of N and m.
+for( N <-  Ns; m <- ms ) {
+
+  val estf = estfactory(m,N)
+    
+    val estname = "simN" + N.toString + "m" + m
+    print("Running " + estname)
+    val eststarttime = (new java.util.Date).getTime
+    
+    //for all the noise distributions (in parallel threads)
+    val mselist = noises.par.map { noise =>
+
+      val siggen =  new PolynomialPhaseSignal(N) //construct a signal generator
+      siggen.setNoiseGenerator(noise)
+      val est = estf() //construct an estimator
+				  
+      var mse = new Array[Double](m+1) //storage for the mses
+      for( itr <- 1 to iters ) {
+	siggen.generateRandomParameters(m)
+	val p0 = siggen.getParameters
+	siggen.generateReceivedSignal
+	val err = est.error(siggen.getReal, siggen.getImag, p0)
+	for( i <- mse.indices ) mse(i) += err(i) 
+      }
+      print(".")
+      mse //last thing is what gets returned 
+    }.toList
+    
+    //now write all the data to a file
+    val files = (0 to m).map( v => new java.io.FileWriter(estname + "p" + v) ) //list of files to write to
+    (mselist, snrdbs).zipped.foreach{ (mse, snrdb) =>
[…]

### 2012-12-06 build.sh
moved more stuff

@@ -1,6 +1,6 @@
-cd code 
+cd code/data
 mpost -interaction=nonstopmode plot.mp 
-cd .. 
+cd ../..
 
 pdflatex paper.tex
 bibtex paper.aux

### 2012-12-06 code/sim.scala
edited sim.scala to write to the data directory

@@ -52,7 +52,7 @@ for( N <-  Ns; m <- ms ) {
     }.toList
     
     //now write all the data to a file
-    val files = (0 to m).map( v => new java.io.FileWriter(estname + "p" + v) ) //list of files to write to
+    val files = (0 to m).map( v => new java.io.FileWriter("data/" + estname + "p" + v) ) //list of files to write to
     (mselist, snrdbs).zipped.foreach{ (mse, snrdb) =>
       for ( i <- files.indices ) 
 	files(i).write(snrdb.toString.replace('E', 'e') + "\t" + (fourpisqr*mse(i)/iters).toString.replace('E', 'e')  + "\n") 

### 2013-02-08 paper.tex
mved stuff
added to elsev

--- 2012-10-19 paper.tex
@@
-%\documentclass[draftcls, onecolumn, 11pt]{../../bib/IEEEtran}
-\documentclass[journal,10pt]{IEEEtran}
+%\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[review]{elsarticle}
+\documentclass[5p]{elsarticle}
 
@@
 
+\journal{Signal Processing}
+
 \begin{document}
+
+\begin{frontmatter}
 
@@
 
-\author{Robby McKilliam and Andr\'{e} Pollok%
-\thanks{Supported under the Australian Government's Australian Space Research Program.
-Robby McKilliam and Andr\'{e} Pollok are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095}
-}
-% The paper headers
-\markboth{Robby McKilliam and Andre Pollock, On the Cram\'{e}r-Rao bound for polynomial phase signals}%
-{DRAFT \today}
-
-% make the title area
-\maketitle
+\author{Robby McKilliam} 
+\author{Andr\'{e} Pollok}
+\address{Institute for Telecommunications Research, The University of South Australia, SA, 5095}
+\fntext[asrp]{Supported under the Australian Government's Australian Space Research Program.}
 
@@
 Of practical importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
-%Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.
-The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.
-We derive closed form formulae for the Cram\'{e}r-Rao  lower bounds of these parameters under the assumption that the noise is additive white and Gaussian.
-This is achieved by making use of a family of orthogonal polynomials.
+Assuming that the noise is additive and Gaussian, the direct evaluation of the Cram\'{e}r-Rao lower bound for this estimation problem involves evaluating the inverse of a matrix.
+Computing this inverse is numerically difficult for polynomial phase signal of large order.
+By making use of a family of orthogonal polynomials, we derive formulae for the Cram\'{e}r-Rao lower bounds that are numerically stable and easy to compute.
 \end{abstract}
 
-\begin{IEEEkeywords}
+\begin{keyword}
 Polynomial-phase signals, Cram\'{e}r-Rao bound.
-\end{IEEEkeywords}
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-%\bibliographystyle{IEEEbib}
-\small 
+\bibliographystyle{elsarticle-num}
+%\small 
 \bibliography{bib}

### 2013-02-08 paper.tex

--- 2012-12-06 paper.tex
@@
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal,10pt]{IEEEtran}
+%\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[review]{elsarticle}
+\documentclass[5p]{elsarticle}
 
@@
 Of practical importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
-%Assuming that the noise is additive and Gaussian,  Peleg and Porat approximated the Cram\'{e}r-Rao bound for this estimation problem.
-The approximation is accurate when the number of observations is sufficiently larger than the order of the polynomial phase signal.
-We derive closed form and numerically stable formulae for the Cram\'{e}r-Rao  lower bounds of these parameters under the assumption that the noise is additive white and Gaussian.
-This is achieved by making use of a family of orthogonal polynomials.
+Assuming that the noise is additive and Gaussian, the direct evaluation of the Cram\'{e}r-Rao lower bound for this estimation problem involves evaluating the inverse of a matrix.
+Computing this inverse is numerically difficult for polynomial phase signal of large order.
+By making use of a family of orthogonal polynomials, we derive formulae for the Cram\'{e}r-Rao lower bounds that are numerically stable and easy to compute.
 \end{abstract}
 
-\begin{IEEEkeywords}
+\begin{keyword}
 Polynomial-phase signals, Cram\'{e}r-Rao bound.
-\end{IEEEkeywords}
+\end{keyword}
+
+\end{frontmatter}
 
@@
 This leads directly to closed form expressions for the Cram\'{e}r-Rao bounds.
-In Section~\ref{sec:stab-cons} we manipulate these expression so as to make their numerical stability clear.
+%In Section~\ref{sec:stab-cons} we manipulate these expression so as to make their numerical stability clear.
 This simultaneously indicates the prefered stable approach to implementing the formula in a digital computer.
@@
 However, including the amplitude and noise variance is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
-Section~\ref{sec:simuations} SIMS BLERG.
+Section~\ref{sec:simuations} %SIMS BLERG.
 

### 2013-02-08 paper.tex
added elsev

@@
 \begin{abstract}
-Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication.
+Polynomial-phase signals have applications including radar, sonar, biology, and radio communication.
 Of practical importance is the estimation of the parameters of a polynomial phase signal from a sequence of noisy observations.
@@
 
-A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model~\cite{Peleg_DPT_1995, Peleg1991_CRB_PPS_1991, Peleg1991_est_class_PPS_1991},
+A uniformly sampled, constant amplitude, polynomial-phase signal of order $m$ has the model,
 \begin{equation}\label{eq:polyestfunction}
@@
 
-Polynomial-phase signals have applications including radar, sonar, geophysics, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001,Hlawatsch_lin_quad_time_freq_spmag_1992}.
-The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
+Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001,Hlawatsch_lin_quad_time_freq_spmag_1992}.
+%The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
 Of practical importance is the estimation of the coefficients $\beta = (\beta_0, \dots \beta_m)\in\reals^{m+1}$ from $N$ consecutive samples $y_1, \dots, y_N$ where 
@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Peleg_DPT_1995, Angeby_estimating_2000, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Barbarossa_analysis_of_PPS_1997, Slocumb_polynomial_1994, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994,Francos_2dpolyest_1998,McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
+Many estimators have been studied and implemented~\cite{Angeby_estimating_2000,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994}.
 
@@
 Here we remove the need for this approximation.
-By converting the polynomial basis to one that is orthogonal we derive a closed form and numerically stable formula for the inverse Fisher information matrix.
-This leads to closed form and numerically stable formulae for the Cram\'{e}r-Rao bounds.
+By converting the polynomial basis to one that is orthogonal we derive a closed form formula for the inverse Fisher information matrix.
+This leads to closed form formulae for the Cram\'{e}r-Rao bounds.
 %At the same time we tie together some existing literature on the Cram\'{e}r-Roa bound for polynomial phase signals~\cite{Peleg1991_CRB_PPS_1991,Ristic_crbpelegcomment_1998}.
@@
 However, including the amplitude and noise variance is straightforward, and we do this in Section~\ref{sec:including-amplitude}.
-Section~\ref{sec:simuations} %SIMS BLERG.
+%Section~\ref{sec:simuations} %SIMS BLERG.
 
@@
 \]  
-The Fisher information matrix $F$ corresponding with $\beta$ is the $m+1$ by $m+1$ matrix with elements $F_{i k} = f_{ik}(\beta)$.  The elements are
+The Fisher information matrix $F$ corresponding with $\beta$ is the $m+1$ by $m+1$ matrix with elements
 \[
-F_{ik} = F_{ki} = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} n^{i}n^{k}.
+F_{ik} = F_{ki} = f_{ik}(\beta) = \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} n^{i}n^{k}.
 \]
@@
 We will circumvent this problem by applying a change of basis to the polynomial coefficients $\beta$.
-This will lead to a closed form and numerically stable formula for $F^{-1}$.
+%This will lead to a closed form and numerically stable formula for $F^{-1}$.
 

### 2013-02-09 paper.tex
minor

@@
 
-Observe that the polynomials $q_0,q_1,\dots$ satisfy the recursion
+\newcommand{\stirling}[2]{\genfrac{[}{]}{0pt}{}{#1}{#2}}
+
+% Observe that the polynomials $q_0,q_1,\dots$ satisfy the recursion
+%  \[
+%  q_s(x) = \frac{x-s}{s} q_{s-1}(x), \qquad q_{0}(x) = 1.
+%  \]
+%  If $q_s(x) = q_{s0} + q_{s1}x + \dots, q_{ss}x^s$ so that $q_{s0},q_{s1},\dots,q_{ss}$ are the coefficients of $q_{s}$, then
+%  \[
+%  q_{si} = \frac{1}{s}q_{s-1,i-1} - q_{s-1,i}
+%  \]
+%  where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if any of $s$ or $i$ is negative.
+
+The $i$th coefficient of $d_k$ is
+\begin{align*}
+P_{ki} = \sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_{si}},
+%&= \sum_{s=0}^k{\frac{(-1)^{k-i}}{s!}\binom{s+k}{s}\binom{N-s-1}{N-k-1} \stirling{s+1}{k+1} }.
+\end{align*}
+where $q_{si}$ is the $i$th coefficient of the polynomial $q_s$.  Observe that
 \[
-q_s(x) = \frac{x-s}{s} q_{s-1}(x), \qquad q_{0}(x) = 1.
+q_{si} = \frac{(-1)^{s-i}}{s!}\stirling{s+1}{k+1}
 \]
-If $q_s(x) = q_{s0} + q_{s1}x + \dots, q_{ss}x^s$ so that $q_{s0},q_{s1},\dots,q_{ss}$ are the coefficients of $q_{s}$, then
-\[
-q_{si} = \frac{1}{s}q_{s-1,i-1} - q_{s-1,i}
-\]
-where $q_{00}=1$ and $q_{0i}=0$ for $i \neq 0$ and $q_{si} = 0$ if any of $s$ or $i$ is negative.
-
-Now the $i$th coefficient of $d_k$ is
-\[
-P_{ki} = \sum_{s=0}^k{(-1)^{s+k}\binom{s+k}{s}\binom{N-s-1}{N-k-1}q_{si}}
-\]  
-which is closed form in the sense that the number of operations required to compute $P_{ki}$ does not depend on $N$.
-The inverse matrix $F^{-1} = P^\prime H^{-1} P$ can now be computed in closed form.
-Evaluating the diagonal terms gives bounds,
+where $\stirling{s}{k}$ denotes the unsigned Stirling number of the first kind~\cite{Abramowitz_Stegen_stirling_numbers_1972}.  
+The inverse matrix $F^{-1} = P^\prime H^{-1} P$ can now be computed in closed form.  Evaluating the diagonal terms gives bounds,
 \begin{align*}
@@
 
-% \section{Simulations}\label{sec:simuations}
+\section{Examples and simulations}\label{sec:simuations}
 
-% Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} display the results of Monte-Carlo simulation with the maximum likelihood estimator and also the exact Cram\'{e}r-Rao bound and the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.
+Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} display the results of Monte-Carlo simulation with the maximum likelihood estimator and also the exact Cram\'{e}r-Rao bound and the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.
 The simulations use $m=2$ and $N=5,20$ and with $\text{SNR} = \tfrac{\alpha^2}{2\sigma^2}$ in the range \unit[-5]{dB} to \unit[25]{dB}.
@@
 
+\begin{figure}[p]
+   	\centering 
+  		\includegraphics{code/data/plot-0.mps} 
+   		\caption{Variance versus SNR.} 
+   		\label{plot:phase0} 
+\end{figure}
+
 % \begin{figure}[p]
 %    	\centering 
-%   		\includegraphics{code/plot-1.mps} 
+%   		\includegraphics{code/data/plot-1.mps} 
 %    		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
@@
 %    	\centering 
-%   		\includegraphics{code/plot-2.mps} 
+%   		\includegraphics{code/data/plot-2.mps} 
 %    		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
@@
 %    	\centering 
-%   		\includegraphics{code/plot-3.mps} 
+%   		\includegraphics{code/data/plot-3.mps} 
 %    		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 

### 2013-03-13 paper.tex
added sig proc version

@@
-%\documentclass[preprint,10pt]{elsarticle}
+\documentclass[preprint,10pt]{elsarticle}
 %\documentclass[review]{elsarticle}
-\documentclass[5p]{elsarticle}
+%\documentclass[5p]{elsarticle}
 
@@
 Computing this inverse is numerically difficult for polynomial phase signal of large order.
-By making use of a family of orthogonal polynomials, we derive formulae for the Cram\'{e}r-Rao lower bounds that are numerically stable and easy to compute.
+By making use of a family of orthogonal polynomials, we derive formulae for the Cram\'{e}r-Rao bounds that are numerically stable and easy to compute.
 \end{abstract}
@@
 
-Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001,Hlawatsch_lin_quad_time_freq_spmag_1992}.
+Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001}.
 %The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Angeby_estimating_2000,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990, Oshea_iterative_1996, Morelande_bayes_unwrapping_2008, Kitchen_polyphase_unwrapping_1994}.
+Many estimators have been studied and implemented~\cite{Oshea_cpf_2004,Morelande_bayes_unwrapping_2009_tsp,Kitchen_polyphase_unwrapping_1994,Angeby_estimating_2000,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990}.
 
-We remark on our notation.
+%We remark on our notation.
 We use lower case Latin letters for working variables and we use Greek letters for fixed numbers.
@@
 If $\hat{\beta}$ is an estimator of $\beta$ then the corresponding estimator of $\gamma$ is $\hat{\gamma} = \hat{\beta}P^{-1}$.
-If $C$ is the covariance of $\hat{\beta}$ then the covariance of $\hat{\gamma}$ is $(P^{-1})^\prime C P^{-1}$ where $^\prime$ is the transpose.
+If $C$ is the covariance matrix of $\hat{\beta}$ then the covariance matrix of $\hat{\gamma}$ is $(P^{-1})^\prime C P^{-1}$ where $^\prime$ indicates transpose.
 
@@
 \] 
-and we can compute the corresponding Fisher information matrix.  It is found to be the $m+1$ by $m+1$ matrix $H$ with elements
+and we can derive the corresponding Fisher information matrix.  It is found to be the $m+1$ by $m+1$ matrix $H$ with elements
 \[
@@
 where the argument $b=(b_0, \dots, b_m)$ represents the polynomial phase coefficients and the argument $a$ represents the amplitude.
-We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$ and $\alpha$ and $\sigma^2$  from observations $y_1, \dots, y_N$ where $y_n = s_n(\beta,\alpha) + w_n$.
+We now consider estimation of $\beta = (\beta_0, \dots, \beta_m)$ and $\alpha$ and $\sigma^2$  from observations $y_n = s_n(\beta,\alpha) + w_n, n = 1,\dots,N$.
 The likelihood function is now
@@
 \]
-where the argument $v$ represents the noise variance.
+where the argument $v$ represents the (hypothesised) noise variance.
 
@@
 The simulations use $m=2$ and $N=5,20$ and with $\text{SNR} = \tfrac{\alpha^2}{2\sigma^2}$ in the range \unit[-5]{dB} to \unit[25]{dB}.
-The average square error of the maximum likelihood estimator after $10^4$ replacations is given by the dots.
+The average square error of the maximum likelihood estimator after $10^4$ replications is given by the dots.
 This is computed using the approach from~\cite[Sec.~10.1]{McKilliam2010thesis}.
 The solid line is the exact Cram\'{e}r-Rao bound and the dashed line is the approximation.
-As is expected, the maximum likelihood estimator achieves the Cram\'{e}r-Rao bound provided that the SNR is large enough, so that the `threshold' effect is avoided.
+The maximum likelihood estimator achieves the Cram\'{e}r-Rao bound provided that the SNR is large enough, so that the `threshold' effect is avoided.
 The approximate bound from~\cite{Peleg1991_CRB_PPS_1991} becomes accurate as $N$ increases.
@@
 
+Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} represent a somewhat artificial situation, since when $m=2$ the $3 \times 3$ matrix $H$ in not poorly conditioned, and can be accurately inverted.
+Our explanation for choosing $m=2$ in Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} is that it is not computationally feasible to compute the maximum likelihood estimator for $m>2$.
+In practice, the matrix $H$ becomes difficult to invert when $m>5$.
+For example, MATLAB warns of numerical imprecision when attempting to invert $H$ when $m>5$.
+Figure~\ref{plot:relerror} displays the relative error of the approximate Cram\'{e}r-Rao bound defined as
+\[
+\text{relative error} = \frac{\abs{\text{CRB}_{\text{approx}} - \text{CRB}}}{\text{CRB}}
+\] 
+for polynomial phase signals of order $m=8$.
+The relative error is larger for the lower order coefficients.
+The relative error decreases as $N$ increases, but is still significant for the low order coefficients when $N=100$.
+
+\section{Conclusion}
+
+Using a family of discrete orthogonal polynomials we describe closed form formulae for the Cram\'{e}r-Rao lower bounds for unbiased estimators of polynomial phase signals.
+
+\bibliographystyle{elsarticle-num}
+\small 
+\bibliography{bib}
+
+% \begin{figure}[p]
+%    	\centering 
+%   		\includegraphics{code/data/plot-0.mps} 
+%    		\caption{Variance versus SNR.} 
+%    		\label{plot:phase0} 
+% \end{figure}
+
  \begin{figure}[p]
     	\centering 
-  		\includegraphics{code/data/plot-0.mps} 
-   		\caption{Variance versus SNR.} 
+   		\includegraphics{code/data/plot-1.mps} 
+    		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
     		\label{plot:phase0} 
@@
 
-% \begin{figure}[p]
-%    	\centering 
-%   		\includegraphics{code/data/plot-1.mps} 
-%    		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
-%    		\label{plot:phase0} 
-% \end{figure}
+ \begin{figure}[p]
+    	\centering 
+   		\includegraphics{code/data/plot-2.mps} 
+    		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
+    		\label{plot:phase1} 
+ \end{figure}
 
-% \begin{figure}[p]
-%    	\centering 
-%   		\includegraphics{code/data/plot-2.mps} 
-%    		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
-%    		\label{plot:phase1} 
-% \end{figure}
+ \begin{figure}[p]
+    	\centering 
+   		\includegraphics{code/data/plot-3.mps} 
+    		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
+    		\label{plot:phase2} 
+ \end{figure}
 
-% \begin{figure}[p]
-%    	\centering 
-%   		\includegraphics{code/data/plot-3.mps} 
-%    		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
-%    		\label{plot:phase2} 
-% \end{figure}
-
-\section{Conclusion}
-
-Using a family of discrete orthogonal polynomials we provide closed form formulae for the Cram\'{e}r-Rao lower bounds for unbiased estimators of polynomial phase signals.
-
-\bibliographystyle{elsarticle-num}
-%\small 
-\bibliography{bib}
+ \begin{figure}[p]
+    	\centering 
+   		\includegraphics{code/data/plot-4.mps} 
+    		\caption{Relative error of approximate Cram\'{e}r-Rao bound from~\cite{Peleg1991_CRB_PPS_1991} when $m=8$.} 
+    		\label{plot:relerror} 
+ \end{figure}
 

### 2013-03-13 paper.tex
put figures on individual pages

@@
-\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[review]{elsarticle}
+%\documentclass[preprint,10pt]{elsarticle}
+\documentclass[review]{elsarticle}
 %\documentclass[5p]{elsarticle}
@@
 
-Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004,gershman_estimating_2001}.
+Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004}.
 %The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
@@
 
+\clearpage
+
  \begin{figure}[p]
@@
  \end{figure}
+
+\clearpage
 

### 2013-03-18 paper.tex
testing url
minor mods

@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Oshea_cpf_2004,Morelande_bayes_unwrapping_2009_tsp,Kitchen_polyphase_unwrapping_1994,Angeby_estimating_2000,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990}.
+Many estimators have been studied and implemented~\cite{Angeby_estimating_2000,Oshea_cpf_2004,Morelande_bayes_unwrapping_2009_tsp,Kitchen_polyphase_unwrapping_1994,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990}.
 
@@
 
-Put,
+Put
 \[
@@
 \begin{align*}
-H_{ik} &= \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} d_{i}(n) d_{k}(n) \\
-&= \begin{cases}
+H_{ik} &= \frac{\alpha^2}{\sigma^2}\sum_{n=1}^{N} d_{i}(n) d_{k}(n) 
+= \begin{cases}
 0 &  k\neq i \\
@@
 \[
-L(b,a,v) = \frac{1}{(2 \pi v )^N} \exp\left( - \frac{1}{2 v } \sum_{n=1}^N \abs{y_n -  s_n(b,a) }^2 \right).
+L(b,a,v) = \frac{1}{(2 \pi v )^N} \exp\left( - \frac{1}{2 v } \sum_{n=1}^N \abs{y_n -  s_n(b,a) }^2 \right),
 \]
@@
 For example, MATLAB warns of numerical imprecision when attempting to invert $H$ when $m>5$.
-Figure~\ref{plot:relerror} displays the relative error of the approximate Cram\'{e}r-Rao bound defined as
-\[
-\text{relative error} = \frac{\abs{\text{CRB}_{\text{approx}} - \text{CRB}}}{\text{CRB}}
-\] 
-for polynomial phase signals of order $m=8$.
+Figure~\ref{plot:relerror} displays the relative error of the approximate Cram\'{e}r-Rao bound for polynomial phase signals of order $m=8$.
+The relative error is defined as $(\tilde{F}_{ii}^{-1} - F_{ii}^{-1})/F_{ii}^{-1}$ where $\tilde{F}_{ii}^{-1}$ is the approximate Cram\'{e}r-Rao bound given by equation (34) in~\cite{Peleg1991_CRB_PPS_1991}.
 The relative error is larger for the lower order coefficients.

### 2013-03-22 build.sh
added pdf figures to the sigproc branch

@@ -1,6 +1,6 @@
-cd code/data
-mpost -interaction=nonstopmode plot.mp 
-cd ../..
+#cd code/data
+#mpost -interaction=nonstopmode plot.mp 
+#cd ../..
 
 pdflatex paper.tex
 bibtex paper.aux

### 2013-03-27 paper.tex
added pdf figures to the sigproc branch
added email addresses to front page
added phone and fax
cleaned

@@
 
-\usepackage{mathbf-abbrevs}
+\newcommand{\abf}{\mathbf{a}}
+\newcommand{\bbf}{\mathbf{b}}
+\newcommand{\cbf}{\mathbf{c}}
+\newcommand{\dbf}{\mathbf{d}}
+\newcommand{\ebf}{\mathbf{e}}
+\newcommand{\fbf}{\mathbf{f}}
+\newcommand{\gbf}{\mathbf{g}}
+\newcommand{\hbf}{\mathbf{h}}
+\newcommand{\ibf}{\mathbf{i}}
+\newcommand{\jbf}{\mathbf{j}}
+\newcommand{\kbf}{\mathbf{k}}
+\newcommand{\lbf}{\mathbf{l}}
+\newcommand{\mbf}{\mathbf{m}}
+\newcommand{\nbf}{\mathbf{n}}
+\newcommand{\obf}{\mathbf{o}}
+\newcommand{\pbf}{\mathbf{p}}
+\newcommand{\qbf}{\mathbf{q}}
+\newcommand{\rbf}{\mathbf{r}}
+\newcommand{\sbf}{\mathbf{s}}
+\newcommand{\tbf}{\mathbf{t}}
+\newcommand{\ubf}{\mathbf{u}}
+\newcommand{\vbf}{\mathbf{v}}
+\newcommand{\wbf}{\mathbf{w}}
+\newcommand{\xbf}{\mathbf{x}}
+\newcommand{\ybf}{\mathbf{y}}
+\newcommand{\zbf}{\mathbf{z}}
+\newcommand{\Abf}{\mathbf{A}}
+\newcommand{\Bbf}{\mathbf{B}}
+\newcommand{\Cbf}{\mathbf{C}}
+\newcommand{\Dbf}{\mathbf{D}}
+\newcommand{\Ebf}{\mathbf{E}}
+\newcommand{\Fbf}{\mathbf{F}}
+\newcommand{\Gbf}{\mathbf{G}}
+\newcommand{\Hbf}{\mathbf{H}}
+\newcommand{\Ibf}{\mathbf{I}}
+\newcommand{\Jbf}{\mathbf{J}}
+\newcommand{\Kbf}{\mathbf{K}}
+\newcommand{\Lbf}{\mathbf{L}}
+\newcommand{\Mbf}{\mathbf{M}}
+\newcommand{\Nbf}{\mathbf{N}}
+\newcommand{\Obf}{\mathbf{O}}
+\newcommand{\Pbf}{\mathbf{P}}
+\newcommand{\Qbf}{\mathbf{Q}}
+\newcommand{\Rbf}{\mathbf{R}}
+\newcommand{\Sbf}{\mathbf{S}}
+\newcommand{\Tbf}{\mathbf{T}}
+\newcommand{\Ubf}{\mathbf{U}}
+\newcommand{\Vbf}{\mathbf{V}}
+\newcommand{\Wbf}{\mathbf{W}}
+\newcommand{\Xbf}{\mathbf{X}}
+\newcommand{\Ybf}{\mathbf{Y}}
+\newcommand{\Zbf}{\mathbf{Z}}
+\newcommand{\alphabf}{\boldsymbol{\alpha}}
+\newcommand{\betabf}{\boldsymbol{\beta}}
+\newcommand{\gammabf}{\boldsymbol{\gamma}}
+\newcommand{\deltabf}{\boldsymbol{\delta}}
+\newcommand{\epsilonbf}{\boldsymbol{\epsilon}}
+\newcommand{\varepsilonbf}{\boldsymbol{\varepsilon}}
+\newcommand{\zetabf}{\boldsymbol{\zeta}}
+\newcommand{\etabf}{\boldsymbol{\eta}}
+\newcommand{\thetabf}{\boldsymbol{\theta}}
+\newcommand{\varthetabf}{\boldsymbol{\vartheta}}
+\newcommand{\iotabf}{\boldsymbol{\iota}}
+\newcommand{\kappabf}{\boldsymbol{\kappa}}
+\newcommand{\lambdabf}{\boldsymbol{\lambda}}
+\newcommand{\mubf}{\boldsymbol{\mu}}
+\newcommand{\nubf}{\boldsymbol{\nu}}
+\newcommand{\xibf}{\boldsymbol{\xi}}
+\newcommand{\pibf}{\boldsymbol{\pi}}
+\newcommand{\varpibf}{\boldsymbol{\varpi}}
+\newcommand{\rhobf}{\boldsymbol{\rho}}
+\newcommand{\varrhobf}{\boldsymbol{\varrho}}
+\newcommand{\sigmabf}{\boldsymbol{\sigma}}
+\newcommand{\varsigmabf}{\boldsymbol{\varsigma}}
+\newcommand{\taubf}{\boldsymbol{\tau}}
+\newcommand{\upsilonbf}{\boldsymbol{\upsilon}}
+\newcommand{\phibf}{\boldsymbol{\phi}}
+\newcommand{\varphibf}{\boldsymbol{\varphi}}
+\newcommand{\chibf}{\boldsymbol{\chi}}
+\newcommand{\psibf}{\boldsymbol{\psi}}
+\newcommand{\omegabf}{\boldsymbol{\omega}}
+\newcommand{\Gammabf}{\boldsymbol{\Gamma}}
+\newcommand{\Deltabf}{\boldsymbol{\Delta}}
+\newcommand{\Thetabf}{\boldsymbol{\Theta}}
+\newcommand{\Lambdabf}{\boldsymbol{\Lambda}}
+\newcommand{\Xibf}{\boldsymbol{\Xi}}
+\newcommand{\Pibf}{\boldsymbol{\Pi}}
+\newcommand{\Sigmabf}{\boldsymbol{\Sigma}}
+\newcommand{\Upsilonbf}{\boldsymbol{\Upsilon}}
+\newcommand{\Phibf}{\boldsymbol{\Phi}}
+\newcommand{\Psibf}{\boldsymbol{\Psi}}
+\newcommand{\Omegabf}{\boldsymbol{\Omega}}
+\newcommand{\zerobf}{\boldsymbol{0}}
+\newcommand{\onebf}{\boldsymbol{1}}
 
-\input{defs}
+\newcommand{\acal}{\mathcal{a}}
+\newcommand{\bcal}{\mathcal{b}}
+\newcommand{\ccal}{\mathcal{c}}
+\newcommand{\dcal}{\mathcal{d}}
+\newcommand{\ecal}{\mathcal{e}}
+\newcommand{\fcal}{\mathcal{f}}
+\newcommand{\gcal}{\mathcal{g}}
+\newcommand{\hcal}{\mathcal{h}}
+\newcommand{\ical}{\mathcal{i}}
+\newcommand{\jcal}{\mathcal{j}}
+\newcommand{\kcal}{\mathcal{k}}
+\newcommand{\lcal}{\mathcal{l}}
+\newcommand{\mcal}{\mathcal{m}}
+\newcommand{\ncal}{\mathcal{n}}
+\newcommand{\ocal}{\mathcal{o}}
+\newcommand{\pcal}{\mathcal{p}}
+\newcommand{\qcal}{\mathcal{q}}
+\newcommand{\rcal}{\mathcal{r}}
+\newcommand{\scal}{\mathcal{s}}
+\newcommand{\tcal}{\mathcal{t}}
+\newcommand{\ucal}{\mathcal{u}}
+\newcommand{\vcal}{\mathcal{v}}
+\newcommand{\wcal}{\mathcal{w}}
+\newcommand{\xcal}{\mathcal{x}}
+\newcommand{\ycal}{\mathcal{y}}
+\newcommand{\zcal}{\mathcal{z}}
+\newcommand{\Acal}{\mathcal{A}}
+\newcommand{\Bcal}{\mathcal{B}}
+\newcommand{\Ccal}{\mathcal{C}}
+\newcommand{\Dcal}{\mathcal{D}}
+\newcommand{\Ecal}{\mathcal{E}}
+\newcommand{\Fcal}{\mathcal{F}}
+\newcommand{\Gcal}{\mathcal{G}}
+\newcommand{\Hcal}{\mathcal{H}}
+\newcommand{\Ical}{\mathcal{I}}
+\newcommand{\Jcal}{\mathcal{J}}
+\newcommand{\Kcal}{\mathcal{K}}
+\newcommand{\Lcal}{\mathcal{L}}
+\newcommand{\Mcal}{\mathcal{M}}
+\newcommand{\Ncal}{\mathcal{N}}
+\newcommand{\Ocal}{\mathcal{O}}
+\newcommand{\Pcal}{\mathcal{P}}
+\newcommand{\Qcal}{\mathcal{Q}}
+\newcommand{\Rcal}{\mathcal{R}}
+\newcommand{\Scal}{\mathcal{S}}
+\newcommand{\Tcal}{\mathcal{T}}
+\newcommand{\Ucal}{\mathcal{U}}
+\newcommand{\Vcal}{\mathcal{V}}
+\newcommand{\Wcal}{\mathcal{W}}
+\newcommand{\Xcal}{\mathcal{X}}
+\newcommand{\Ycal}{\mathcal{Y}}
+\newcommand{\Zcal}{\mathcal{Z}}
+
+\newcommand{\abb}{\mathbb{a}}
+\newcommand{\bbb}{\mathbb{b}}
+\newcommand{\cbb}{\mathbb{c}}
+\newcommand{\dbb}{\mathbb{d}}
+\newcommand{\ebb}{\mathbb{e}}
+\newcommand{\fbb}{\mathbb{f}}
+\newcommand{\gbb}{\mathbb{g}}
+\newcommand{\hbb}{\mathbb{h}}
+\newcommand{\ibb}{\mathbb{i}}
+\newcommand{\jbb}{\mathbb{j}}
+\newcommand{\kbb}{\mathbb{k}}
+\newcommand{\lbb}{\mathbb{l}}
+\newcommand{\mbb}{\mathbb{m}}
+\newcommand{\nbb}{\mathbb{n}}
+\newcommand{\obb}{\mathbb{o}}
+\newcommand{\pbb}{\mathbb{p}}
+\newcommand{\qbb}{\mathbb{q}}
+\newcommand{\rbb}{\mathbb{r}}
+\newcommand{\sbb}{\mathbb{s}}
+\newcommand{\tbb}{\mathbb{t}}
+\newcommand{\ubb}{\mathbb{u}}
+\newcommand{\vbb}{\mathbb{v}}
+\newcommand{\wbb}{\mathbb{w}}
+\newcommand{\xbb}{\mathbb{x}}
+\newcommand{\ybb}{\mathbb{y}}
+\newcommand{\zbb}{\mathbb{z}}
+\newcommand{\Abb}{\mathbb{A}}
+\newcommand{\Bblb}{\mathbb{B}}
+\newcommand{\Cbb}{\mathbb{C}}
+\newcommand{\Dbb}{\mathbb{D}}
+\newcommand{\Ebb}{\mathbb{E}}
+\newcommand{\Fbb}{\mathbb{F}}
+\newcommand{\Gbb}{\mathbb{G}}
+\newcommand{\Hbb}{\mathbb{H}}
+\newcommand{\Ibb}{\mathbb{I}}
+\newcommand{\Jbb}{\mathbb{J}}
+\newcommand{\Kbb}{\mathbb{K}}
+\newcommand{\Lbb}{\mathbb{L}}
+\newcommand{\Mbb}{\mathbb{M}}
+\newcommand{\Nbb}{\mathbb{N}}
+\newcommand{\Obb}{\mathbb{O}}
+\newcommand{\Pbb}{\mathbb{P}}
+\newcommand{\Qbb}{\mathbb{Q}}
+\newcommand{\Rbb}{\mathbb{R}}
+\newcommand{\Sbb}{\mathbb{S}}
+\newcommand{\Tbb}{\mathbb{T}}
+\newcommand{\Ubb}{\mathbb{U}}
+\newcommand{\Vbb}{\mathbb{V}}
+\newcommand{\Wbb}{\mathbb{W}}
+\newcommand{\Xbb}{\mathbb{X}}
+\newcommand{\Ybb}{\mathbb{Y}}
+\newcommand{\Zbb}{\mathbb{Z}}
+
+\newcommand{\mathhbf}[1]{\hat{\mathbf{#1}}}
+\newcommand{\ahbf}{\mathhbf{a}}
+\newcommand{\bhbf}{\mathhbf{b}}
+\newcommand{\chbf}{\mathhbf{c}}
+\newcommand{\dhbf}{\mathhbf{d}}
+\newcommand{\ehbf}{\mathhbf{e}}
+\newcommand{\fhbf}{\mathhbf{f}}
+\newcommand{\ghbf}{\mathhbf{g}}
+\newcommand{\hhbf}{\mathhbf{h}}
+\newcommand{\ihbf}{\mathhbf{i}}
+\newcommand{\jhbf}{\mathhbf{j}}
+\newcommand{\khbf}{\mathhbf{k}}
+\newcommand{\lhbf}{\mathhbf{l}}
+\newcommand{\mhbf}{\mathhbf{m}}
+\newcommand{\nhbf}{\mathhbf{n}}
+\newcommand{\ohbf}{\mathhbf{o}}
+\newcommand{\phbf}{\mathhbf{p}}
+\newcommand{\qhbf}{\mathhbf{q}}
+\newcommand{\rhbf}{\mathhbf{r}}
+\newcommand{\shbf}{\mathhbf{s}}
+\newcommand{\thbf}{\mathhbf{t}}
+\newcommand{\uhbf}{\mathhbf{u}}
+\newcommand{\vhbf}{\mathhbf{v}}
+\newcommand{\whbf}{\mathhbf{w}}
+\newcommand{\xhbf}{\mathhbf{x}}
+\newcommand{\yhbf}{\mathhbf{y}}
+\newcommand{\zhbf}{\mathhbf{z}}
+\newcommand{\Ahbf}{\mathhbf{A}}
+\newcommand{\Bhbf}{\mathhbf{B}}
+\newcommand{\Chbf}{\mathhbf{C}}
+\newcommand{\Dhbf}{\mathhbf{D}}
+\newcommand{\Ehbf}{\mathhbf{E}}
+\newcommand{\Fhbf}{\mathhbf{F}}
+\newcommand{\Ghbf}{\mathhbf{G}}
+\newcommand{\Hhbf}{\mathhbf{H}}
+\newcommand{\Ihbf}{\mathhbf{I}}
+\newcommand{\Jhbf}{\mathhbf{J}}
+\newcommand{\Khbf}{\mathhbf{K}}
+\newcommand{\Lhbf}{\mathhbf{L}}
+\newcommand{\Mhbf}{\mathhbf{M}}
+\newcommand{\Nhbf}{\mathhbf{N}}
+\newcommand{\Ohbf}{\mathhbf{O}}
+\newcommand{\Phbf}{\mathhbf{P}}
+\newcommand{\Qhbf}{\mathhbf{Q}}
+\newcommand{\Rhbf}{\mathhbf{R}}
+\newcommand{\Shbf}{\mathhbf{S}}
+\newcommand{\Thbf}{\mathhbf{T}}
+\newcommand{\Uhbf}{\mathhbf{U}}
+\newcommand{\Vhbf}{\mathhbf{V}}
+\newcommand{\Whbf}{\mathhbf{W}}
+\newcommand{\Xhbf}{\mathhbf{X}}
+\newcommand{\Yhbf}{\mathhbf{Y}}
+\newcommand{\Zhbf}{\mathhbf{Z}}
+
+\newcommand{\ah}{\hat{a}}
+\newcommand{\bh}{\hat{b}}
+\newcommand{\ch}{\hat{c}}
+\newcommand{\dhat}{\hat{d}}
+\newcommand{\eh}{\hat{e}}
+\newcommand{\fh}{\hat{f}}
+\newcommand{\gh}{\hat{g}}
+\newcommand{\hh}{\hat{h}}
+\newcommand{\ih}{\hat{i}}
+\newcommand{\jh}{\hat{j}}
+\newcommand{\kh}{\hat{k}}
+\newcommand{\lh}{\hat{l}}
+\newcommand{\mh}{\hat{m}}
+\newcommand{\nh}{\hat{n}}
+\newcommand{\oh}{\hat{o}}
+\newcommand{\ph}{\hat{p}}
+\newcommand{\qh}{\hat{q}}
+\newcommand{\rh}{\hat{r}}
+\newcommand{\sh}{\hat{s}}
+\newcommand{\that}{\hat{t}}
+\newcommand{\uh}{\hat{u}}
+\newcommand{\vh}{\hat{v}}
+\newcommand{\wh}{\hat{w}}
+\newcommand{\xh}{\hat{x}}
+\newcommand{\yh}{\hat{y}}
+\newcommand{\zh}{\hat{z}}
+\newcommand{\Ah}{\hat{A}}
+\newcommand{\Bh}{\hat{B}}
+\newcommand{\Ch}{\hat{C}}
+\newcommand{\Dh}{\hat{D}}
+\newcommand{\Eh}{\hat{E}}
+\newcommand{\Fh}{\hat{F}}
+\newcommand{\Gh}{\hat{G}}
+\newcommand{\Hh}{\hat{H}}
+\newcommand{\Ih}{\hat{I}}
+\newcommand{\Jh}{\hat{J}}
+\newcommand{\Kh}{\hat{K}}
+\newcommand{\Lh}{\hat{L}}
+\newcommand{\Mh}{\hat{M}}
+\newcommand{\Nh}{\hat{N}}
+\newcommand{\Oh}{\hat{O}}
+\newcommand{\Ph}{\hat{P}}
+\newcommand{\Qh}{\hat{Q}}
+\newcommand{\Rh}{\hat{R}}
+\newcommand{\Sh}{\hat{S}}
+\newcommand{\Th}{\hat{T}}
+\newcommand{\Uh}{\hat{U}}
+\newcommand{\Vh}{\hat{V}}
+\newcommand{\Wh}{\hat{W}}
+\newcommand{\Xh}{\hat{X}}
+\newcommand{\Yh}{\hat{Y}}
+\newcommand{\Zh}{\hat{Z}} 
+\newcommand{\thetah}{\hat{\theta}} 
+\newcommand{\phih}{\hat{\phi}}
+
+\newcommand{\atbf}{\tilde{\mathbf{a}}}
+\newcommand{\btbf}{\tilde{\mathbf{b}}}
+\newcommand{\ctbf}{\tilde{\mathbf{c}}}
+\newcommand{\dtbf}{\tilde{\mathbf{d}}}
+\newcommand{\etbf}{\tilde{\mathbf{e}}}
+\newcommand{\ftbf}{\tilde{\mathbf{f}}}
+\newcommand{\gtbf}{\tilde{\mathbf{g}}}
+\newcommand{\htbf}{\tilde{\mathbf{h}}}
+\newcommand{\itbf}{\tilde{\mathbf{i}}}
+\newcommand{\jtbf}{\tilde{\mathbf{j}}}
+\newcommand{\ktbf}{\tilde{\mathbf{k}}}
+\newcommand{\ltbf}{\tilde{\mathbf{l}}}
+\newcommand{\mtbf}{\tilde{\mathbf{m}}}
+\newcommand{\ntbf}{\tilde{\mathbf{n}}}
+\newcommand{\otbf}{\tilde{\mathbf{o}}}
+\newcommand{\ptbf}{\tilde{\mathbf{p}}}
+\newcommand{\qtbf}{\tilde{\mathbf{q}}}
+\newcommand{\rtbf}{\tilde{\mathbf{r}}}
+\newcommand{\stbf}{\tilde{\mathbf{s}}}
+\newcommand{\ttbf}{\tilde{\mathbf{t}}}
+\newcommand{\utbf}{\tilde{\mathbf{u}}}
+\newcommand{\vtbf}{\tilde{\mathbf{v}}}
+\newcommand{\wtbf}{\tilde{\mathbf{w}}}
+\newcommand{\xtbf}{\tilde{\mathbf{x}}}
+\newcommand{\ytbf}{\tilde{\mathbf{y}}}
+\newcommand{\ztbf}{\tilde{\mathbf{z}}}
+\newcommand{\Atbf}{\tilde{\mathbf{A}}}
+\newcommand{\Btbf}{\tilde{\mathbf{B}}}
+\newcommand{\Ctbf}{\tilde{\mathbf{C}}}
+\newcommand{\Dtbf}{\tilde{\mathbf{D}}}
+\newcommand{\Etbf}{\tilde{\mathbf{E}}}
+\newcommand{\Ftbf}{\tilde{\mathbf{F}}}
+\newcommand{\Gtbf}{\tilde{\mathbf{G}}}
+\newcommand{\Htbf}{\tilde{\mathbf{H}}}
+\newcommand{\Itbf}{\tilde{\mathbf{I}}}
+\newcommand{\Jtbf}{\tilde{\mathbf{J}}}
+\newcommand{\Ktbf}{\tilde{\mathbf{K}}}
+\newcommand{\Ltbf}{\tilde{\mathbf{L}}}
+\newcommand{\Mtbf}{\tilde{\mathbf{M}}}
+\newcommand{\Ntbf}{\tilde{\mathbf{N}}}
+\newcommand{\Otbf}{\tilde{\mathbf{O}}}
+\newcommand{\Ptbf}{\tilde{\mathbf{P}}}
+\newcommand{\Qtbf}{\tilde{\mathbf{Q}}}
+\newcommand{\Rtbf}{\tilde{\mathbf{R}}}
+\newcommand{\Stbf}{\tilde{\mathbf{S}}}
+\newcommand{\Ttbf}{\tilde{\mathbf{T}}}
+\newcommand{\Utbf}{\tilde{\mathbf{U}}}
+\newcommand{\Vtbf}{\tilde{\mathbf{V}}}
+\newcommand{\Wtbf}{\tilde{\mathbf{W}}}
+\newcommand{\Xtbf}{\tilde{\mathbf{X}}}
+\newcommand{\Ytbf}{\tilde{\mathbf{Y}}}
+\newcommand{\Ztbf}{\tilde{\mathbf{Z}}}
+\newcommand{\alphatbf}{\tilde{\boldsymbol{\alpha}}}
+\newcommand{\betatbf}{\tilde{\boldsymbol{\beta}}}
+\newcommand{\gammatbf}{\tilde{\boldsymbol{\gamma}}}
+\newcommand{\deltatbf}{\tilde{\boldsymbol{\delta}}}
+\newcommand{\epsilontbf}{\tilde{\boldsymbol{\epsilon}}}
+\newcommand{\varepsilontbf}{\tilde{\boldsymbol{\varepsilon}}}
+\newcommand{\zetatbf}{\tilde{\boldsymbol{\zeta}}}
+\newcommand{\etatbf}{\tilde{\boldsymbol{\eta}}}
+\newcommand{\thetatbf}{\tilde{\boldsymbol{\theta}}}
+\newcommand{\varthetatbf}{\tilde{\boldsymbol{\vartheta}}}
+\newcommand{\iotatbf}{\tilde{\boldsymbol{\iota}}}
+\newcommand{\kappatbf}{\tilde{\boldsymbol{\kappa}}}
+\newcommand{\lambdatbf}{\tilde{\boldsymbol{\lambda}}}
+\newcommand{\mutbf}{\tilde{\boldsymbol{\mu}}}
+\newcommand{\nutbf}{\tilde{\boldsymbol{\nu}}}
+\newcommand{\xitbf}{\tilde{\boldsymbol{\xi}}}
+\newcommand{\pitbf}{\tilde{\boldsymbol{\pi}}}
+\newcommand{\varpitbf}{\tilde{\boldsymbol{\varpi}}}
+\newcommand{\rhotbf}{\tilde{\boldsymbol{\rho}}}
+\newcommand{\varrhotbf}{\tilde{\boldsymbol{\varrho}}}
+\newcommand{\sigmatbf}{\tilde{\boldsymbol{\sigma}}}
+\newcommand{\varsigmatbf}{\tilde{\boldsymbol{\varsigma}}}
+\newcommand{\tautbf}{\tilde{\boldsymbol{\tau}}}
+\newcommand{\upsilontbf}{\tilde{\boldsymbol{\upsilon}}}
+\newcommand{\phitbf}{\tilde{\boldsymbol{\phi}}}
+\newcommand{\varphitbf}{\tilde{\boldsymbol{\varphi}}}
+\newcommand{\chitbf}{\tilde{\boldsymbol{\chi}}}
+\newcommand{\psitbf}{\tilde{\boldsymbol{\psi}}}
+\newcommand{\omegatbf}{\tilde{\boldsymbol{\omega}}}
+\newcommand{\Gammatbf}{\tilde{\boldsymbol{\Gamma}}}
+\newcommand{\Deltatbf}{\tilde{\boldsymbol{\Delta}}}
+\newcommand{\Thetatbf}{\tilde{\boldsymbol{\Theta}}}
+\newcommand{\Lambdatbf}{\tilde{\boldsymbol{\Lambda}}}
+\newcommand{\Xitbf}{\tilde{\boldsymbol{\Xi}}}
+\newcommand{\Pitbf}{\tilde{\boldsymbol{\Pi}}}
+\newcommand{\Sigmatbf}{\tilde{\boldsymbol{\Sigma}}}
+\newcommand{\Upsilontbf}{\tilde{\boldsymbol{\Upsilon}}}
+\newcommand{\Phitbf}{\tilde{\boldsymbol{\Phi}}}
+\newcommand{\Psitbf}{\tilde{\boldsymbol{\Psi}}}
+\newcommand{\Omegatbf}{\tilde{\boldsymbol{\Omega}}}
+\newcommand{\zerotbf}{\tilde{\boldsymbol{0}}}
+\newcommand{\onetbf}{\tilde{\boldsymbol{1}}}
+
+\newcommand{\at}{\tilde{a}}
+\newcommand{\bt}{\tilde{b}}
+\newcommand{\ct}{\tilde{c}}
+\newcommand{\dt}{\tilde{d}}
+\newcommand{\et}{\tilde{e}}
+\newcommand{\ft}{\tilde{f}}
+\newcommand{\gt}{\tilde{g}}
+\newcommand{\hti}{\tilde{h}}
+\newcommand{\iti}{\tilde{i}}
+\newcommand{\jt}{\tilde{j}}
+\newcommand{\kt}{\tilde{k}}
+\newcommand{\lt}{\tilde{l}}
+\newcommand{\mt}{\tilde{m}}
+\newcommand{\nt}{\tilde{n}}
+\newcommand{\ot}{\tilde{o}}
+\newcommand{\pt}{\tilde{p}}
+\newcommand{\qt}{\tilde{q}}
+\newcommand{\rt}{\tilde{r}}
+\newcommand{\st}{\tilde{s}}
+\newcommand{\tti}{\tilde{t}}
+\newcommand{\ut}{\tilde{u}}
+\newcommand{\vt}{\tilde{v}}
+\newcommand{\wt}{\tilde{w}}
+\newcommand{\xt}{\tilde{x}}
+\newcommand{\yt}{\tilde{y}}
+\newcommand{\zt}{\tilde{z}}
+\newcommand{\At}{\tilde{A}}
+\newcommand{\Bt}{\tilde{B}}
+\newcommand{\Ct}{\tilde{C}}
+\newcommand{\Dt}{\tilde{D}}
+\newcommand{\Et}{\tilde{E}}
+\newcommand{\Ft}{\tilde{F}}
+\newcommand{\Gt}{\tilde{G}}
+\newcommand{\Ht}{\tilde{H}}
+\newcommand{\It}{\tilde{I}}
+\newcommand{\Jt}{\tilde{J}}
+\newcommand{\Kt}{\tilde{K}}
+\newcommand{\Lt}{\tilde{L}}
+\newcommand{\Mt}{\tilde{M}}
+\newcommand{\Nt}{\tilde{N}}
+\newcommand{\Ot}{\tilde{O}}
+\newcommand{\Pt}{\tilde{P}}
+\newcommand{\Qt}{\tilde{Q}}
+\newcommand{\Rt}{\tilde{R}}
+\newcommand{\St}{\tilde{S}}
+\newcommand{\Tt}{\tilde{T}}
+\newcommand{\Ut}{\tilde{U}}
+\newcommand{\Vt}{\tilde{V}}
+\newcommand{\Wt}{\tilde{W}}
+\newcommand{\Xt}{\tilde{X}}
+\newcommand{\Yt}{\tilde{Y}}
+\newcommand{\Zt}{\tilde{Z}} 
+\newcommand{\phit}{\tilde{\phi}}
+
+\newcommand{\aubbf}{\mathbf{\underline{a}}}
+\newcommand{\bubbf}{\mathbf{\underline{b}}}
+\newcommand{\cubbf}{\mathbf{\underline{c}}}
+\newcommand{\dubbf}{\mathbf{\underline{d}}}
+\newcommand{\eubbf}{\mathbf{\underline{e}}}
+\newcommand{\fubbf}{\mathbf{\underline{f}}}
+\newcommand{\gubbf}{\mathbf{\underline{g}}}
+\newcommand{\hubbf}{\mathbf{\underline{h}}}
+\newcommand{\iubbf}{\mathbf{\underline{i}}}
+\newcommand{\jubbf}{\mathbf{\underline{j}}}
+\newcommand{\kubbf}{\mathbf{\underline{k}}}
+\newcommand{\lubbf}{\mathbf{\underline{l}}}
+\newcommand{\mubbf}{\mathbf{\underline{m}}}
+\newcommand{\nubbf}{\mathbf{\underline{n}}}
+\newcommand{\oubbf}{\mathbf{\underline{o}}}
+\newcommand{\pubbf}{\mathbf{\underline{p}}}
+\newcommand{\qubbf}{\mathbf{\underline{q}}}
+\newcommand{\rubbf}{\mathbf{\underline{r}}}
+\newcommand{\subbf}{\mathbf{\underline{s}}}
+\newcommand{\tubbf}{\mathbf{\underline{t}}}
+\newcommand{\uubbf}{\mathbf{\underline{u}}}
+\newcommand{\vubbf}{\mathbf{\underline{v}}}
+\newcommand{\wubbf}{\mathbf{\underline{w}}}
+\newcommand{\xubbf}{\mathbf{\underline{x}}}
+\newcommand{\yubbf}{\mathbf{\underline{y}}}
+\newcommand{\zubbf}{\mathbf{\underline{z}}}
+\newcommand{\Aubbf}{\mathbf{\underline{A}}}
+\newcommand{\Bubbf}{\mathbf{\underline{B}}}
+\newcommand{\Cubbf}{\mathbf{\underline{C}}}
+\newcommand{\Dubbf}{\mathbf{\underline{D}}}
+\newcommand{\Eubbf}{\mathbf{\underline{E}}}
+\newcommand{\Fubbf}{\mathbf{\underline{F}}}
+\newcommand{\Gubbf}{\mathbf{\underline{G}}}
+\newcommand{\Hubbf}{\mathbf{\underline{H}}}
+\newcommand{\Iubbf}{\mathbf{\underline{I}}}
+\newcommand{\Jubbf}{\mathbf{\underline{J}}}
+\newcommand{\Kubbf}{\mathbf{\underline{K}}}
+\newcommand{\Lubbf}{\mathbf{\underline{L}}}
+\newcommand{\Mubbf}{\mathbf{\underline{M}}}
+\newcommand{\Nubbf}{\mathbf{\underline{N}}}
+\newcommand{\Oubbf}{\mathbf{\underline{O}}}
+\newcommand{\Pubbf}{\mathbf{\underline{P}}}
+\newcommand{\Qubbf}{\mathbf{\underline{Q}}}
+\newcommand{\Rubbf}{\mathbf{\underline{R}}}
+\newcommand{\Subbf}{\mathbf{\underline{S}}}
+\newcommand{\Tubbf}{\mathbf{\underline{T}}}
+\newcommand{\Uubbf}{\mathbf{\underline{U}}}
+\newcommand{\Vubbf}{\mathbf{\underline{V}}}
+\newcommand{\Wubbf}{\mathbf{\underline{W}}}
+\newcommand{\Xubbf}{\mathbf{\underline{X}}}
+\newcommand{\Yubbf}{\mathbf{\underline{Y}}}
+\newcommand{\Zubbf}{\mathbf{\underline{Z}}}
+\newcommand{\alphaubbf}{\underline{\boldsymbol{\alpha}}}
+\newcommand{\betaubbf}{\underline{\boldsymbol{\beta}}}
+\newcommand{\gammaubbf}{\underline{\boldsymbol{\gamma}}}
+\newcommand{\deltaubbf}{\underline{\boldsymbol{\delta}}}
+\newcommand{\epsilonubbf}{\underline{\boldsymbol{\epsilon}}}
+\newcommand{\varepsilonubbf}{\underline{\boldsymbol{\varepsilon}}}
+\newcommand{\zetaubbf}{\underline{\boldsymbol{\zeta}}}
+\newcommand{\etaubbf}{\underline{\boldsymbol{\eta}}}
+\newcommand{\thetaubbf}{\underline{\boldsymbol{\theta}}}
+\newcommand{\varthetaubbf}{\underline{\boldsymbol{\vartheta}}}
+\newcommand{\iotaubbf}{\underline{\boldsymbol{\iota}}}
+\newcommand{\kappaubbf}{\underline{\boldsymbol{\kappa}}}
+\newcommand{\lambdaubbf}{\underline{\boldsymbol{\lambda}}}
+\newcommand{\muubbf}{\underline{\boldsymbol{\mu}}}
+\newcommand{\nuubbf}{\underline{\boldsymbol{\nu}}}
+\newcommand{\xiubbf}{\underline{\boldsymbol{\xi}}}
+\newcommand{\piubbf}{\underline{\boldsymbol{\pi}}}
+\newcommand{\varpiubbf}{\underline{\boldsymbol{\varpi}}}
+\newcommand{\rhoubbf}{\underline{\boldsymbol{\rho}}}
+\newcommand{\varrhoubbf}{\underline{\boldsymbol{\varrho}}}
+\newcommand{\sigmaubbf}{\underline{\boldsymbol{\sigma}}}
+\newcommand{\varsigmaubbf}{\underline{\boldsymbol{\varsigma}}}
+\newcommand{\tauubbf}{\underline{\boldsymbol{\tau}}}
+\newcommand{\upsilonubbf}{\underline{\boldsymbol{\upsilon}}}
+\newcommand{\phiubbf}{\underline{\boldsymbol{\phi}}}
+\newcommand{\varphiubbf}{\underline{\boldsymbol{\varphi}}}
+\newcommand{\chiubbf}{\underline{\boldsymbol{\chi}}}
+\newcommand{\psiubbf}{\underline{\boldsymbol{\psi}}}
+\newcommand{\omegaubbf}{\underline{\boldsymbol{\omega}}}
+\newcommand{\Gammaubbf}{\underline{\boldsymbol{\Gamma}}}
+\newcommand{\Deltaubbf}{\underline{\boldsymbol{\Delta}}}
+\newcommand{\Thetaubbf}{\underline{\boldsymbol{\Theta}}}
+\newcommand{\Lambdaubbf}{\underline{\boldsymbol{\Lambda}}}
+\newcommand{\Xiubbf}{\underline{\boldsymbol{\Xi}}}
+\newcommand{\Piubbf}{\underline{\boldsymbol{\Pi}}}
+\newcommand{\Sigmaubbf}{\underline{\boldsymbol{\Sigma}}}
+\newcommand{\Upsilonubbf}{\underline{\boldsymbol{\Upsilon}}}
+\newcommand{\Phiubbf}{\underline{\boldsymbol{\Phi}}}
+\newcommand{\Psiubbf}{\underline{\boldsymbol{\Psi}}}
+\newcommand{\Omegaubbf}{\underline{\boldsymbol{\Omega}}}
+\newcommand{\zeroubbf}{\underline{\boldsymbol{0}}}
+\newcommand{\oneubbf}{\underline{\boldsymbol{1}}}
+
+\newcommand{\aub}{\underline{a}}
+\newcommand{\bub}{\underline{b}}
+\newcommand{\cub}{\underline{c}}
+\newcommand{\dub}{\underline{d}}
+\newcommand{\eub}{\underline{e}}
+\newcommand{\fub}{\underline{f}}
+\newcommand{\gub}{\underline{g}}
+\newcommand{\hub}{\underline{h}}
+\newcommand{\iub}{\underline{i}}
+\newcommand{\jub}{\underline{j}}
+\newcommand{\kub}{\underline{k}}
+\newcommand{\lub}{\underline{l}}
+\newcommand{\mub}{\underline{m}}
+\newcommand{\nub}{\underline{n}}
+\newcommand{\oub}{\underline{o}}
+\newcommand{\pub}{\underline{p}}
+\newcommand{\qub}{\underline{q}}
+\newcommand{\rub}{\underline{r}}
+\newcommand{\sub}{\underline{s}}
+\newcommand{\tub}{\underline{t}}
+\newcommand{\uub}{\underline{u}}
+\newcommand{\vub}{\underline{v}}
+\newcommand{\wub}{\underline{w}}
+\newcommand{\xub}{\underline{x}}
+\newcommand{\yub}{\underline{y}}
+\newcommand{\zub}{\underline{z}}
+\newcommand{\Aub}{\underline{A}}
+\newcommand{\Bub}{\underline{B}}
+\newcommand{\Cub}{\underline{C}}
+\newcommand{\Dub}{\underline{D}}
+\newcommand{\Eub}{\underline{E}}
+\newcommand{\Fub}{\underline{F}}
+\newcommand{\Gub}{\underline{G}}
+\newcommand{\Hub}{\underline{H}}
+\newcommand{\Iub}{\underline{I}}
+\newcommand{\Jub}{\underline{J}}
+\newcommand{\Kub}{\underline{K}}
+\newcommand{\Lub}{\underline{L}}
+\newcommand{\Mub}{\underline{M}}
+\newcommand{\Nub}{\underline{N}}
+\newcommand{\Oub}{\underline{O}}
+\newcommand{\Pub}{\underline{P}}
+\newcommand{\Qub}{\underline{Q}}
+\newcommand{\Rub}{\underline{R}}
+\newcommand{\Sub}{\underline{S}}
+\newcommand{\Tub}{\underline{T}}
+\newcommand{\Uub}{\underline{U}}
+\newcommand{\Vub}{\underline{V}}
+\newcommand{\Wub}{\underline{W}}
+\newcommand{\Xub}{\underline{X}}
+\newcommand{\Yub}{\underline{Y}}
+\newcommand{\Zub}{\underline{Z}}
+\newcommand{\alphaub}{\underline{\alpha}}
+\newcommand{\betaub}{\underline{\beta}}
+\newcommand{\gammaub}{\underline{\gamma}}
+\newcommand{\deltaub}{\underline{\delta}}
+\newcommand{\epsilonub}{\underline{\epsilon}}
+\newcommand{\varepsilonub}{\underline{\varepsilon}}
+\newcommand{\zetaub}{\underline{\zeta}}
+\newcommand{\etaub}{\underline{\eta}}
+\newcommand{\thetaub}{\underline{\theta}}
+\newcommand{\varthetaub}{\underline{\vartheta}}
+\newcommand{\iotaub}{\underline{\iota}}
+\newcommand{\kappaub}{\underline{\kappa}}
+\newcommand{\lambdaub}{\underline{\lambda}}
+\newcommand{\muub}{\underline{\mu}}
+\newcommand{\nuub}{\underline{\nu}}
+\newcommand{\xiub}{\underline{\xi}}
+\newcommand{\piub}{\underline{\pi}}
+\newcommand{\varpiub}{\underline{\varpi}}
+\newcommand{\rhoub}{\underline{\rho}}
+\newcommand{\varrhoub}{\underline{\varrho}}
+\newcommand{\sigmaub}{\underline{\sigma}}
+\newcommand{\varsigmaub}{\underline{\varsigma}}
+\newcommand{\tauub}{\underline{\tau}}
+\newcommand{\upsilonub}{\underline{\upsilon}}
+\newcommand{\phiub}{\underline{\phi}}
+\newcommand{\varphiub}{\underline{\varphi}}
+\newcommand{\chiub}{\underline{\chi}}
+\newcommand{\psiub}{\underline{\psi}}
+\newcommand{\omegaub}{\underline{\omega}}
+\newcommand{\Gammaub}{\underline{\Gamma}}
+\newcommand{\Deltaub}{\underline{\Delta}}
+\newcommand{\Thetaub}{\underline{\Theta}}
+\newcommand{\Lambdaub}{\underline{\Lambda}}
+\newcommand{\Xiub}{\underline{\Xi}}
+\newcommand{\Piub}{\underline{\Pi}}
+\newcommand{\Sigmaub}{\underline{\Sigma}}
+\newcommand{\Upsilonub}{\underline{\Upsilon}}
+\newcommand{\Phiub}{\underline{\Phi}}
+\newcommand{\Psiub}{\underline{\Psi}}
+\newcommand{\Omegaub}{\underline{\Omega}}
+\newcommand{\zeroub}{\underline{0}}
+\newcommand{\oneub}{\underline{1}}
+
+\newcommand{\xbfML}{\xbf_{\text{ML}}}
+\newcommand{\hbfML}{\hbf_{\text{ML}}}
+\newcommand{\XbfML}{\Xbf_{\text{ML}}}
+\newcommand{\HbfML}{\Hbf_{\text{ML}}}
+
+%\newcommand {\tbf}[1] {\textbf{#1}}
+%\newcommand {\tit}[1] {\textit{#1}}
+%\newcommand {\tmd}[1] {\textmd{#1}}
+%\newcommand {\trm}[1] {\textrm{#1}}
+%\newcommand {\tsc}[1] {\textsc{#1}}
+%\newcommand {\tsf}[1] {\textsf{#1}}
+%\newcommand {\tsl}[1] {\textsl{#1}}
+%\newcommand {\ttt}[1] {\texttt{#1}}
+%\newcommand {\tup}[1] {\textup{#1}}
+%
+%\newcommand {\mbf}[1] {\mathbf{#1}}
+%\newcommand {\mmd}[1] {\mathmd{#1}}
+%\newcommand {\mrm}[1] {\mathrm{#1}}
+%\newcommand {\msc}[1] {\mathsc{#1}}
+%\newcommand {\msf}[1] {\mathsf{#1}}
+%\newcommand {\msl}[1] {\mathsl{#1}}
+%\newcommand {\mtt}[1] {\mathtt{#1}}
+%\newcommand {\mup}[1] {\mathup{#1}}
+
+%some math functions and symbols
+\newcommand{\reals}{{\mathbb R}}
+\newcommand{\ints}{{\mathbb Z}}
+\newcommand{\integers}{{\mathbb Z}}
+\newcommand{\sign}[1]{\mathtt{sign}\left( #1 \right)}
+\newcommand{\NP}{\operatorname{NearestPt}}
+\newcommand{\NS}{\operatorname{NearestSet}}
+\newcommand{\bres}{\operatorname{Bres}}
+\newcommand{\vol}{\operatorname{vol}}
+\newcommand{\vor}{\operatorname{vor}}
+\newcommand{\coef}{\operatorname{coef}}
+\newcommand{\Int}{\operatorname{Int}}
+
+\newcommand{\selectindicies}{\operatorname{selectindices}}
+\newcommand{\sortindicies}{\operatorname{sortindices}}
+\newcommand{\largest}{\operatorname{largest}}
+\newcommand{\quickpartition}{\operatorname{quickpartition}}
+\newcommand{\quickpartitiontwo}{\operatorname{quickpartition2}}
+
+%some commonly used underlined and
+%hated symbols
+\newcommand{\uY}{\ushort{\mbf{Y}}}
+\newcommand{\ueY}{\ushort{Y}}
+\newcommand{\uy}{\ushort{\mbf{y}}}
+\newcommand{\uey}{\ushort{y}}
+\newcommand{\ux}{\ushort{\mbf{x}}}
+\newcommand{\uex}{\ushort{x}}
+\newcommand{\uhx}{\ushort{\mbf{\hat{x}}}}
+\newcommand{\uehx}{\ushort{\hat{x}}}
+
+\newcommand{\term}{\emph}
+\newcommand{\var}{\operatorname{var}}
+\newcommand{\covar}{\operatorname{covar}}
+%\newcommand{\prob}{{\mathbb P}}
+\newcommand{\prob}{\operatorname{Pr}}
+\newcommand{\expect}{{\mathbb E}}
+\newcommand{\dealias}{\operatorname{dealias}}
+\renewcommand{\mid}{\; ; \;}
+
+% Brackets
+\newcommand{\br}[1]{{\left( #1 \right)}}
+\newcommand{\sqbr}[1]{{\left[ #1 \right]}}
+\newcommand{\cubr}[1]{{\left\{ #1 \right\}}}
+\newcommand{\abr}[1]{\left< #1 \right>}
+\newcommand{\abs}[1]{{\left| #1 \right|}}
+\newcommand{\floor}[1]{{\left\lfloor #1 \right\rfloor}}
+\newcommand{\ceiling}[1]{{\left\lceil #1 \right\rceil}}
+\newcommand{\ceil}[1]{\lceil #1 \rceil}
+\newcommand{\round}[1]{{\left\lceil #1 \right\rfloor}}
+\newcommand{\magn}[1]{\left\| #1 \right\|}
+\newcommand{\fracpart}[1]{\left\{ #1 \right\}}
+
+% Referencing
+\newcommand{\refeqn}[1]{\eqref{#1}}
+\newcommand{\reffig}[1]{Figure~\ref{#1}}
+\newcommand{\reftable}[1]{Table~\ref{#1}}
+\newcommand{\refsec}[1]{Section~\ref{#1}}
+\newcommand{\refappendix}[1]{Appendix~\ref{#1}}
+\newcommand{\refchapter}[1]{Chapter~\ref{#1}}
+
+\newcommand {\figwidth} {100mm}
+\newcommand {\Ref}[1] {Reference~\cite{#1}}
+\newcommand {\Sec}[1] {Section~\ref{#1}}
+\newcommand {\App}[1] {Appendix~\ref{#1}}
+\newcommand {\Chap}[1] {Chapter~\ref{#1}}
+\newcommand {\Lem}[1] {Lemma~\ref{#1}}
+\newcommand {\Thm}[1] {Theorem~\ref{#1}}
+\newcommand {\Cor}[1] {Corollary~\ref{#1}}
+\newcommand {\Alg}[1] {Algorithm~\ref{#1}}
+\newcommand {\etal} {\emph{~et~al.}}
+\newcommand {\bul} {$\bullet$ }   % bullet
+\newcommand {\fig}[1] {Figure~\ref{#1}}   % references Figure x
+\newcommand {\imp} {$\Rightarrow$}   % implication symbol (default)
+\newcommand {\impt} {$\Rightarrow$}   % implication symbol (text mode)
+\newcommand {\impm} {\Rightarrow}   % implication symbol (math mode)
+\newcommand {\vect}[1] {\mathbf{#1}} 
+\newcommand {\hvect}[1] {\hat{\mathbf{#1}}}
+\newcommand {\del} {\partial}
+\newcommand {\eqn}[1] {Equation~(\ref{#1})} 
+\newcommand {\tab}[1] {Table~\ref{#1}} % references Table x
+\newcommand {\half} {\frac{1}{2}} 
+\newcommand {\ten}[1] {\times10^{#1}}
+\newcommand {\bra}[2] {\mbox{}_{#2}\langle #1 |}
+\newcommand {\ket}[2] {| #1 \rangle_{#2}}
+\newcommand {\Bra}[2] {\mbox{}_{#2}\left.\left\langle #1 \right.\right|}
+\newcommand {\Ket}[2] {\left.\left| #1 \right.\right\rangle_{#2}}
+\newcommand {\im} {\mathrm{Im}}
+\newcommand {\re} {\mathrm{Re}}
+\newcommand {\braket}[4] {\mbox{}_{#3}\langle #1 | #2 \rangle_{#4}} 
+\newcommand {\trace}[1] {\text{tr}\left(#1\right)}
+\newcommand{\dotprod}[2]{ #1 \cdot #2}
+
+% spell things correctly
+\newenvironment{centre}{\begin{center}}{\end{center}}
+\newenvironment{itemise}{\begin{itemize}}{\end{itemize}}
+
+\usepackage{units}
+	%nice looking units
+
+\usepackage{booktabs}
+		%creates nice looking tables
+
+\usepackage{graphicx}% standard graphics package for inclusion of
+		      % images and eps files into LaTeX document
+
+\usepackage{amsmath,amsfonts,amssymb, amsthm, bm} % this is handy for mathematicians and physicists
+			      % see http://www.ams.org/tex/amslatex.html
+
+\usepackage[vlined, linesnumbered]{algorithm2e}
+	%algorithm writing package
+
+\usepackage{mathrsfs}
+%fancy math script
+
+\usepackage{ushort}
+%enable good underlining in math mode
+
+%------------------------------------------------------------
+% Theorem like environments
+%
+\newtheorem{theorem}{Theorem}
+%\theoremstyle{plain}
+\newtheorem{acknowledgement}{Acknowledgement}
+%\newtheorem{algorithm}{Algorithm}
+\newtheorem{axiom}{Axiom}
+\newtheorem{case}{Case}
+\newtheorem{claim}{Claim}
+\newtheorem{conclusion}{Conclusion}
+\newtheorem{condition}{Condition}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{corollary}{Corollary}
+\newtheorem{criterion}{Criterion}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{exercise}{Exercise}
+\newtheorem{lemma}{Lemma}
+\newtheorem{notation}{Notation}
+\newtheorem{problem}{Problem}
+\newtheorem{proposition}{Proposition}
+\newtheorem{property}{Property}
+\newtheorem{remark}{Remark}
+\newtheorem{solution}{Solution}
+\newtheorem{summary}{Summary}
+%\numberwithin{equation}{section}
+%--------------------------------------------------------
 
@@
 
-\author{Robby McKilliam} 
+\author{Robby McKilliam\corref{cor1}} 
+\ead{robby.mckilliam@unisa.edu.au}
+%\phone{+61 8 830 23602}
+%\fax{+61 8 830 23873}
 \author{Andr\'{e} Pollok}
+%\ead{andre.pollok@unisa.edu.au}
 \address{Institute for Telecommunications Research, The University of South Australia, SA, 5095}
+
+\cortext[cor1]{Corresponding author, Ph: +61 8 830 23602, Fax: +61 8 830 23873}
+
 \fntext[asrp]{Supported under the Australian Government's Australian Space Research Program.}
@@
 \small 
-\bibliography{bib}
+%\bibliography{bib}
+
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{Angeby_estimating_2000}
+J.~\"Angeby, Estimating signal parameters using the nonlinear instantaneous
+  least squares approach, IEEE Trans. Sig. Process. 48 (2000) 2721--2732.
+
+\bibitem{Levanon_Radar_signals_2004}
+N.~Levanon, E.~Mozeson, Radar Signals, John Wiley \& Sons, 2004.
+
+\bibitem{Oshea_cpf_2004}
+P.~O'Shea, A fast algorithm for estimating the parameters of a quadratic {FM}
+  signal, IEEE Trans. Sig. Process. 52~(2) (2004) 385 -- 393.
+
+\bibitem{Morelande_bayes_unwrapping_2009_tsp}
+M.~Morelande, Parameter estimation of phase-modulated signals using {Bayesian}
+  unwrapping, IEEE Trans. Sig. Process. 57~(11) (2009) 4209--4219.
+
+\bibitem{Kitchen_polyphase_unwrapping_1994}
+J.~Kitchen, A method for estimating the coefficients of a polynomial phase
+  signal, Signal Processing 37~(1) (1994) 463--470.
+
+\bibitem{Peleg_DPT_1995}
+S.~Peleg, B.~Friedlander, The discrete polynomial-phase transform, IEEE Trans.
+  Sig. Process. 43~(8) (1995) 1901--1914.
+
+\bibitem{Djuric_phase_unwrap_chirp_1990}
+P.~M. Djuric, S.~M. Kay, Parameter estimation of chirp signals, IEEE Trans.
+  Acoust. Speech Signal Process. 38~(12) (1990) 2118--2126.
+
+\bibitem{Peleg1991_CRB_PPS_1991}
+S.~Peleg, B.~Porat, The {Cram\'{e}r-Rao} lower bound for signals with constant
+  amplitude and polynomial phase, IEEE Trans. Sig. Process. 39~(3) (1991)
+  749--752.
+
+\bibitem{Kay1993_stat_sig_est_theory}
+S.~M. Kay, Fundamentals of Statistical Signal Processing: Estimation Theory,
+  Prentice-Hall, 1993.
+
+\bibitem{Bjork_num_methods_least_square_1996}
+A.~B\"{j}orck, Numerical Methods for Least Squares Problems, SIAM, 1996.
+
+\bibitem{Chebyshev_discrete_polys1864}
+P.~Chebyshev, Sur l'interpolation, Zapiski Akademii Nauk 1~(4) (1864) 539--560.
+
+\bibitem{Gram_discrete_polys1883}
+J.~P. Gram, {\"{U}ber die Entwicklung reeller Funktionen in Reihen mittels der
+  Methode der kleinsten Quadrate}, Journal f\"{u}r die reine und angewandte
+  Mathematik 94 (1883) 41--73.
+
+\bibitem{Eisinberg2007_discerete_otho_poly_equidist}
+A.~Eisinberg, G.~Fedele, Discrete orthogonal polynomials on equidistant nodes,
+  International Mathematical Forum 2~(21) (2007) 1007--1020.
+
+\bibitem{Abramowitz_Stegen_stirling_numbers_1972}
+M.~Abramowitz, I.~A. Stegun, Stirling Numbers of the First Kind, 9th Edition,
+  Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical
+  Tables, Dover, 1972.
+
+\bibitem{McKilliam2010thesis}
+R.~G. McKilliam, Lattice theory, circular statistics and polynomial phase
+  signals, Ph.D. thesis, University of Queensland, Australia (December 2010).
+
+\bibitem{McKilliam2009IndentifiabliltyAliasingPolyphase}
+R.~G. McKilliam, I.~V.~L. Clarkson, Identifiability and aliasing in
+  polynomial-phase signals, IEEE Trans. Sig. Process. 57~(11) (2009)
+  4554--4557.
+
+\end{thebibliography}
 
@@
     	\centering 
-   		\includegraphics{code/data/plot-1.mps} 
+   		\includegraphics{fig1.pdf} 
     		\caption{Variance of $\hat{\beta}_0$ versus SNR.} 
@@
     	\centering 
-   		\includegraphics{code/data/plot-2.mps} 
+   		\includegraphics{fig2.pdf} 
     		\caption{Variance of $\hat{\beta}_1$ versus SNR.} 
@@
     	\centering 
-   		\includegraphics{code/data/plot-3.mps} 
+   		\includegraphics{fig3.pdf} 
     		\caption{Variance of $\hat{\beta}_2$ versus SNR.} 
@@
     	\centering 
-   		\includegraphics{code/data/plot-4.mps} 
+   		\includegraphics{fig4.pdf} 
     		\caption{Relative error of approximate Cram\'{e}r-Rao bound from~\cite{Peleg1991_CRB_PPS_1991} when $m=8$.} 

### 2013-03-27 build.sh

@@ -3,7 +3,6 @@
 #cd ../..
 
 pdflatex paper.tex
-bibtex paper.aux
 pdflatex paper.tex
 pdflatex paper.tex
 

### 2013-03-28 paper.tex
fixed refs

@@
 
-Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000,Levanon_Radar_signals_2004}.
+Polynomial-phase signals have applications including radar, sonar, biology, and radio communication~\cite{Angeby_estimating_2000}.
 %The signals are also used to describe the sounds emitted by bats for echo-location \cite{Peleg_DPT_1995,Moss_2005echolocation,Suga_1975_bats_echolocation}.
@@
 and $\{w_n, n \in \ints\}$ is a sequence of noise variables.
-Many estimators have been studied and implemented~\cite{Angeby_estimating_2000,Oshea_cpf_2004,Morelande_bayes_unwrapping_2009_tsp,Kitchen_polyphase_unwrapping_1994,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990}.
+Many estimators have been studied and implemented~\cite{Angeby_estimating_2000,Oshea_cpf_2004,Morelande_bayes_unwrapping_2009_tsp,Peleg_DPT_1995, Djuric_phase_unwrap_chirp_1990}.
 
@@
 \section{Conclusion}
-
+%\vspace{-0.2cm}
 Using a family of discrete orthogonal polynomials we describe closed form formulae for the Cram\'{e}r-Rao lower bounds for unbiased estimators of polynomial phase signals.
@@
   \def\href#1#2{#2} \def\path#1{#1}\fi
-
+%\vspace{-0.2cm}
 \bibitem{Angeby_estimating_2000}
@@
   least squares approach, IEEE Trans. Sig. Process. 48 (2000) 2721--2732.
-
-\bibitem{Levanon_Radar_signals_2004}
-N.~Levanon, E.~Mozeson, Radar Signals, John Wiley \& Sons, 2004.
-
+%\vspace{-0.1cm}
+%\bibitem{Levanon_Radar_signals_2004}
+%N.~Levanon, E.~Mozeson, Radar Signals, John Wiley \& Sons, 2004.
+%\vspace{-0.1cm}
 \bibitem{Oshea_cpf_2004}
@@
   signal, IEEE Trans. Sig. Process. 52~(2) (2004) 385 -- 393.
-
+%\vspace{-0.1cm}
 \bibitem{Morelande_bayes_unwrapping_2009_tsp}
@@
   unwrapping, IEEE Trans. Sig. Process. 57~(11) (2009) 4209--4219.
-
-\bibitem{Kitchen_polyphase_unwrapping_1994}
-J.~Kitchen, A method for estimating the coefficients of a polynomial phase
-  signal, Signal Processing 37~(1) (1994) 463--470.
-
+%\vspace{-0.1cm}
+%\bibitem{Kitchen_polyphase_unwrapping_1994}
+%J.~Kitchen, A method for estimating the coefficients of a polynomial phase
+%  signal, Signal Processing 37~(1) (1994) 463--470.
+%\vspace{-0.1cm}
 \bibitem{Peleg_DPT_1995}
@@
   Sig. Process. 43~(8) (1995) 1901--1914.
-
+%\vspace{-0.1cm}
 \bibitem{Djuric_phase_unwrap_chirp_1990}
@@
   Acoust. Speech Signal Process. 38~(12) (1990) 2118--2126.
-
+%\vspace{-0.1cm}
 \bibitem{Peleg1991_CRB_PPS_1991}
@@
   749--752.
-
+%\vspace{-0.1cm}
 \bibitem{Kay1993_stat_sig_est_theory}
@@
   Prentice-Hall, 1993.
-
+%\vspace{-0.1cm}
 \bibitem{Bjork_num_methods_least_square_1996}
 A.~B\"{j}orck, Numerical Methods for Least Squares Problems, SIAM, 1996.
-
+%\vspace{-0.1cm}
 \bibitem{Chebyshev_discrete_polys1864}
 P.~Chebyshev, Sur l'interpolation, Zapiski Akademii Nauk 1~(4) (1864) 539--560.
-
+%\vspace{-0.1cm}
 \bibitem{Gram_discrete_polys1883}
@@
   Mathematik 94 (1883) 41--73.
-
+%\vspace{-0.1cm}
 \bibitem{Eisinberg2007_discerete_otho_poly_equidist}
@@
   International Mathematical Forum 2~(21) (2007) 1007--1020.
-
+%\vspace{-0.1cm}
 \bibitem{Abramowitz_Stegen_stirling_numbers_1972}
@@
   Tables, Dover, 1972.
-
+%\vspace{-0.1cm}
 \bibitem{McKilliam2010thesis}
@@
   signals, Ph.D. thesis, University of Queensland, Australia (December 2010).
-
+%\vspace{-0.1cm}
 \bibitem{McKilliam2009IndentifiabliltyAliasingPolyphase}

### 2013-07-05 paper.tex
added to pps

--- 2013-03-18 paper.tex
@@
 Assuming that the noise is additive and Gaussian, the direct evaluation of the Cram\'{e}r-Rao lower bound for this estimation problem involves evaluating the inverse of a matrix.
-Computing this inverse is numerically difficult for polynomial phase signal of large order.
+Computing this inverse is numerically difficult for polynomial phase signals of large order.
 By making use of a family of orthogonal polynomials, we derive formulae for the Cram\'{e}r-Rao bounds that are numerically stable and easy to compute.
@@
 \begin{equation}\label{eq:yndatamodel}
-y_n = s_n(\beta) + w_n
+y_n = s_n(\beta) + w_n \qquad n = 1,\dots, N
 \end{equation}
@@
 
-%We remark on our notation.
-We use lower case Latin letters for working variables and we use Greek letters for fixed numbers.
-Where possible we use the Latin letter and its Greek counterpart when there is a relationship between a working variable and a fixed number.
-For example we use $b=(b_0,\dots,b_m)$ as a working variable to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
-It is common in the literature to use a subscript, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also an element in the vector $b$.
+We remark on our notation.
+We use lowercase letters for working variables and we use Greek letters for fixed parameters.
+Where possible we use the lowercase letter and its Greek counterpart when there is a relationship between a working variable and a fixed parameter.
+For example we use $b=(b_0,\dots,b_m)$ as arguments to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) unknown coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
+It is common in the literature to use a subscript zero, for example $b_0$ to denote the true parameter, but this notation is confusing in our case since $b_0$ is also an element in the vector $b$.
 
@@
 323]{Bjork_num_methods_least_square_1996}\cite{Chebyshev_discrete_polys1864,Gram_discrete_polys1883}.
-The specific family given by Definition~\ref{def:discreteLegendepolys} also appears in~\cite{Eisinberg2007_discerete_otho_poly_equidist}.
+The specific family given by Definition~\ref{def:discreteLegendepolys} appears in~\cite{Eisinberg2007_discerete_otho_poly_equidist}.
 In this paper, we refer to $d_0,d_1,\dots$ simply as the discrete orthogonal polynomials, but the reader should be aware that many other families of polynomials also go by this name.
@@
 The effect is that as SNR decreases, the variance of $\hat{\beta}_k$ converges to that of the uniform distribution on $[-\tfrac{0.5}{k!},\tfrac{0.5}{k!})$.
+A similar effect has been studied in the context of circular statistics~\cite{6365847}.
 
-Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} represent a somewhat artificial situation, since when $m=2$ the $3 \times 3$ matrix $H$ in not poorly conditioned, and can be accurately inverted.
+Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} represent a somewhat artificial situation, since when $m=2$ the $3 \times 3$ matrix $F$ is not poorly conditioned, and can be accurately inverted.
 Our explanation for choosing $m=2$ in Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} is that it is not computationally feasible to compute the maximum likelihood estimator for $m>2$.
-In practice, the matrix $H$ becomes difficult to invert when $m>5$.
-For example, MATLAB warns of numerical imprecision when attempting to invert $H$ when $m>5$.
+In practice, the matrix $F$ becomes difficult to invert when $m>5$.
+For example, MATLAB warns of numerical imprecision when attempting to invert $F$ when $m>5$.
 Figure~\ref{plot:relerror} displays the relative error of the approximate Cram\'{e}r-Rao bound for polynomial phase signals of order $m=8$.
@@
 Using a family of discrete orthogonal polynomials we describe closed form formulae for the Cram\'{e}r-Rao lower bounds for unbiased estimators of polynomial phase signals.
+This avoids direct inversion of the Fisher information matrix, and circumvents the need for the approximate formula given in~\cite{Peleg1991_CRB_PPS_1991}.
 

### 2013-07-09 paper.tex
minor

@@
 \usepackage{mathbf-abbrevs}
+%\usepackage{setspace}
 
@@
 We remark on our notation.
-We use lowercase letters for working variables and we use Greek letters for fixed parameters.
-Where possible we use the lowercase letter and its Greek counterpart when there is a relationship between a working variable and a fixed parameter.
+We use lowercase Roman letters for working variables and we use Greek letters for fixed parameters.
+Where possible we use the lowercase Roman letter and its Greek counterpart when there is a relationship between a working variable and a fixed parameter.
 For example we use $b=(b_0,\dots,b_m)$ as arguments to represent the polynomial phase coefficients in~\eqref{eq:polyestfunction}, and we use $\beta = (\beta_0, \dots \beta_m)$ to denote the fixed (true) unknown coefficients that we desire to estimate in~\eqref{eq:yndatamodel}.
@@
 
-Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} represent a somewhat artificial situation, since when $m=2$ the $3 \times 3$ matrix $F$ is not poorly conditioned, and can be accurately inverted.
+Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} represent a somewhat artificial situation, since when $m=2$ the $3 \times 3$ matrix $F$ is not poorly conditioned, and can be accurately inverted numerically.
 Our explanation for choosing $m=2$ in Figures~\ref{plot:phase0},~\ref{plot:phase1}~and~\ref{plot:phase2} is that it is not computationally feasible to compute the maximum likelihood estimator for $m>2$.
@@
 
+{
 \bibliographystyle{elsarticle-num}
-\small 
+\tiny 
 \bibliography{bib}
+}
 
